### 2013-09-05 paper_draft1.Rnw
Committing existing materials

\documentclass{sigchi}

\pagenumbering{arabic}

<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(fig.path='plots/', fig.align='center', fig.show='asis', echo=FALSE, comment = "", message = FALSE, tidy=FALSE, dev='pdf', out.width="\\linewidth")
options(replace.assign=TRUE)
library(knitr)
library(formatR)
library(ggplot2)
library(lsr)
library(psych)
setwd("~/Dropbox/r-code/SUS/")
sus2 = read.csv("sus2_data.csv")
orisus = read.csv("ori_sus_data.csv")
revsus = read.csv("rev_sus_data.csv")
@

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{SUS 2.0: Updating the System Usability Scale to conform with insights from questionnaire design research}

\numberofauthors{2}
\author{
  \alignauthor Ren\'{e} F. Kizilcec\\
    \affaddr{Department of Communication}\\
    \affaddr{Stanford University}\\
    \email{kizilcec@stanford.edu}
  \alignauthor Hendrik Mueller\\
    \affaddr{UX Research}\\
    \affaddr{Google}\\
    \email{hendrikm@google.com}
}

\maketitle

\begin{abstract}

\end{abstract}

\keywords{}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

\section{Introduction}

<<>>=
# SUS 2.0
# q1 How easy or difficult was it to learn how to use the system?
# q2 How much more is there to learn about the system?
# q3 How confident were you using the system?
# q4 How easy or difficult is it to use the system?
# q5 How consistent is the system?

# Criterion 
cor.test(oe,rowSums(sus2))
cor.test(difficulty,rowSums(sus2))
cor.test(redo,rowSums(sus2))
cor.test(learnsat,rowSums(sus2))

# Internal Consistency
cor.test(rowSums(sus2), q1)
cor.test(rowSums(sus2), q2)
cor.test(rowSums(sus2), q3)
cor.test(rowSums(sus2), q4)
cor.test(rowSums(sus2), q5)

# Cronbach's Alpha
alpha(sus2)

# Split Half Reliability
rhh=cor(rowSums(cbind(q1,q3,q5), na.rm=T), rowSums(cbind(q2,q4), na.rm=T))
(2*rhh)/(1+rhh)

# Factor Analysis
plot(princomp(~q1+q2+q3+q4+q5, data=sus2, scores=T), type="lines")

eigen(cov(sus2, use="pair"))
eigen.loadings(eigen(cov(sus2, use="pair")))
princomp(covmat=cov(sus2, use="pair"))$loadings
plot(princomp(covmat=cov(sus2, use="pair")), type="lines")

fa.parallel(sus2)

# REV SUS
# I needed to learn very few things before I could get going with this system
# I think that I would not need any support of a technical person to be able to use this system
# I did not feel very confident using the system
# I found the system very manageable to use
# I would imagine that most people would learn to use this system very slowly
# I found the system appropriately simple
# I thought the system was hard to use
# I found the various functions in this system were not well integrated
# I did not think there was too much inconsistency in this system
# I do not think that I would like to use this system frequently

alpha(rev_sus)
fa.diagram(fa(rev_sus,1))
fa.diagram(fa(rev_sus,2))

# Original SUS
# I needed to learn a lot of things before I could get going with this system
# I think that I would need the support of a technical person to be able to use this system
# I felt very confident using the system
# I found the system very cumbersome to use
# I would imagine that most people would learn to use this system very quickly
# I found the system unnecessarily complex
# I thought the system was easy to use
# I found the various functions in this system were well integrated
# I thought there was too much inconsistency in this system
# I think that I would like to use this system frequently

alpha(ori_sus)
fa.diagram(fa(ori_sus,1))
fa.diagram(fa(ori_sus,2))

cor.test(oe,rowSums(ori_sus))
cor.test(difficulty,rowSums(ori_sus))
cor.test(redo,rowSums(ori_sus))
cor.test(learnsat,rowSums(ori_sus))

t.test(ori_sus$p1, rev_sus$p1)
t.test(ori_sus$p2, rev_sus$p2)
t.test(ori_sus$p3, rev_sus$p3)
t.test(ori_sus$p4, rev_sus$p4)
t.test(ori_sus$p5, rev_sus$p5) #not
t.test(ori_sus$p6, rev_sus$p6) #not
t.test(ori_sus$p7, rev_sus$p7)
t.test(ori_sus$p8, rev_sus$p8) #not
t.test(ori_sus$p9, rev_sus$p9)
t.test(ori_sus$p10, rev_sus$p10)

t.test(rowMeans(ori_sus), rowMeans(rev_sus))

@

\section{Acknowledgments}

\bibliographystyle{acm-sigchi}
\bibliography{suslit}
\end{document}

### 2013-09-05 prep_data.R

@@ -1,0 +1,143 @@
+library(psych)
+library(plyr)
+setwd("~/Dropbox/r-code/SUS/")
+dat=read.csv("medstats_250813.csv")
+dat=read.csv("medstats_280813.csv")
+names(dat)
+
+extractQ = function(x){
+	temp = as.factor(as.character(x[-1]))
+	return(temp)
+}
+extractQ2 = function(x){
+	temp = as.character(x[-1])
+	return(temp)
+}
+fac2num = function(x){
+	return(as.numeric(as.character(x)))
+}
+
+# Remove duplicates and test responses
+dat = subset(dat, a!="")
+dat = dat[!duplicated(dat$a),]
+
+# Correlates
+oe=extractQ(dat$Q1.1)
+oe=fac2num(factor(oe, labels=c(NA,0,6,2,4,3,1,5)))
+
+difficulty=extractQ(dat$Q1.2)
+difficulty=fac2num(factor(difficulty, labels=c(NA,0,6,3,2,4,1,5)))
+
+redo=extractQ(dat$Q1.3)
+redo=6-fac2num(factor(redo, labels=c(NA,0,6,2,4,3,1,5)))
+
+learnsat=extractQ(dat$Q1.4)
+learnsat=fac2num(factor(learnsat, labels=c(NA,0,6,3,2,4,1,5)))
+
+cor(cbind(oe,difficulty,redo,learnsat), use="pair")
+
+browser=cbind(extractQ2(dat$Q59_1_TEXT), extractQ2(dat$Q76_1_TEXT), extractQ2(dat$Q62_1_TEXT))
+os=cbind(extractQ2(dat$Q59_3_TEXT), extractQ2(dat$Q76_3_TEXT), extractQ2(dat$Q62_3_TEXT))
+res=cbind(extractQ2(dat$Q59_4_TEXT), extractQ2(dat$Q76_4_TEXT), extractQ2(dat$Q62_4_TEXT))
+for(i in 1:nrow(browser)){
+	browser[i,1] = ifelse( sum(browser[i,]!="")==1 , browser[i,(browser[i,]!="")], "")
+	os[i,1] = ifelse( sum(os[i,]!="")==1 , os[i,(os[i,]!="")], "")
+	res[i,1] = ifelse( sum(res[i,]!="")==1 , res[i,(res[i,]!="")], "")
+}
+browser=browser[,1]
+os=os[,1]
+res=res[,1]
+
+# try to get timing which didn't export right
+
[…]

### 2013-09-10 paper_draft1.Rnw
Prepared data from other course and started on paper structure

@@
 library(psych)
-setwd("~/Dropbox/r-code/SUS/")
-sus2 = read.csv("sus2_data.csv")
-orisus = read.csv("ori_sus_data.csv")
-revsus = read.csv("rev_sus_data.csv")
+setwd("~/Dropbox/r-code/SUS2/")
+m.sus2 = read.csv("sus2_medstat.csv")
+m.orisus = read.csv("ori_sus_medstat.csv")
+m.revsus = read.csv("rev_sus_medstat.csv")
+a.sus2 = read.csv("sus2_algo.csv")
+a.orisus = read.csv("ori_sus_algo.csv")
+a.revsus = read.csv("rev_sus_algo.csv")
 @
@@
 \usepackage{graphics} % for EPS, load graphicx instead
-\usepackage{times}    % comment if you want LaTeX's default font
+%\usepackage{times}    % comment if you want LaTeX's default font
 \usepackage{url}      % llt: nicely formatted URLs
@@
 
-<<>>=
+The SUS is used extensively to evaluate systems' usability and it is thus important that it provides accurate results.
+Since its introduction in 19XX, there have been critical advancement in the literature on questionnaire and question design regarding the identification of question biases and how to avoid them.
+As a matter of fact, the SUS is vulnerable to several well-known biases as a result of how the items and scale are constructed.
+
+\section{Background}
+
+Maybe some background on the development and testing of the SUS.
+
+\subsection{Survey Biases}
+\subsubsection{Satisficing}
+\subsubsection{Acquiescence}
+\subsubsection{Question order}
+\subsubsection{Social Desirability}
+\subsubsection{Answer Options}
+\subsubsection{Hypotheticals}
+\subsubsection{Leading Information}
+
+\section{Scale Evaluation}
+
+\subsection{Data}
+Course participants (N$_A$=1746 and N$_B$=406) from two online education courses (denoted A and B) offered by Stanford University were asked to take an end-of-course survey which randomly assigned them into one of three weighted groups: original SUS (25\%), reversed SUS (25\%), and SUS 2.0 (50\%).
+This yielded N$_A^{original}$=439, N$_A^{reversed}$=438, and N$_A^{2.0}$=869 from one course, and N$_B^{original}$=96, N$_B^{reversed}$=105, and N$_B^{2.0}$=205 from the other course.
+The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
+In addition, as part of the survey, respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, their satisfaction with the amount they learnt, and the difficulty of the course.
+
+The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design.
+Based on a design heuristic evaluation of the two systems, it was determined that B had more usability problems than A.
+[Need to talk about how this was evaluated.
+E.g.
+\url{http://en.wikipedia.org/wiki/Heuristic_evaluation}]
+
+\subsection{Acquiescence Bias in the SUS}
+
+<<eval=FALSE>>=
+alpha(m.revsus[,1:10])$total
+alpha(a.revsus[,1:10])$total
+alpha(m.orisus[,1:10])$total
+alpha(a.orisus[,1:10])$total
+
+omega(m.revsus[,1:10])
+omega(a.revsus[,1:10])
+omega(m.orisus[,1:10])
+omega(a.orisus[,1:10])
+
+fa.diagram(fa(m.revsus[,1:10],1))
+fa.diagram(fa(m.revsus[,1:10],2))
+
+head(a.sus2)
+
+# Original SUS
+# I needed to learn a lot of things before I could get going with this system
+# I think that I would need the support of a technical person to be able to use this system
+# I felt very confident using the system
+# I found the system very cumbersome to use
+# I would imagine that most people would learn to use this system very quickly
+# I found the system unnecessarily complex
+# I thought the system was easy to use
+# I found the various functions in this system were well integrated
+# I thought there was too much inconsistency in this system
+# I think that I would like to use this system frequently
+
+alpha(m.orisus[,1:10])
+fa.diagram(fa(ori_sus,1))
+fa.diagram(fa(ori_sus,2))
+
+cor.test(oe,rowSums(ori_sus))
+cor.test(difficulty,rowSums(ori_sus))
+cor.test(redo,rowSums(ori_sus))
+cor.test(learnsat,rowSums(ori_sus))
+
+t.test(ori_sus$p1, rev_sus$p1)
+t.test(ori_sus$p2, rev_sus$p2)
+t.test(ori_sus$p3, rev_sus$p3)
+t.test(ori_sus$p4, rev_sus$p4)
+t.test(ori_sus$p5, rev_sus$p5) #not
+t.test(ori_sus$p6, rev_sus$p6) #not
+t.test(ori_sus$p7, rev_sus$p7)
+t.test(ori_sus$p8, rev_sus$p8) #not
+t.test(ori_sus$p9, rev_sus$p9)
+t.test(ori_sus$p10, rev_sus$p10)
+
+t.test(rowMeans(ori_sus), rowMeans(rev_sus))
+
+@
+
+\subsection{Evaluating the SUS 2.0}
+
+<<eval=FALSE>>=
 # SUS 2.0
@@
 
-\section{Acknowledgments}
+\begin{table*}
+\small
+\centering
+\caption{Items from the original SUS, reversed SUS, and SUS 2.0}
+\label{tab:items}
+\begin{tabular}{| r | p{5cm} | p{5cm} | p{5cm} |}
+\hline
+\#  & Original SUS & Reversed SUS & SUS 2.0 \\
+\hline
+1  & I needed to learn a lot of things before I could get going with this system   & I needed to learn very few things before I could get going with this system & How much more is there to learn about the system?
+\\ \hline
+2  & I think that I would need the support of a technical person to be able to use this system & I think that I would not need any support of a technical person to be able to use this system & How likely are you to need support of a technical person to be able to use the system?
+\\ \hline
+3  & I felt very confident using the system    & I did not feel very confident using the system & How confident are you using the system? \\ \hline
+4  & I found the system very cumbersome to use & I found the system very manageable to use      & How cumbersome is it to use the system? \\ \hline
+5  & I would imagine that most people would learn to use this system very quickly  & I would imagine that most people would learn to use this system very slowly & How easy or difficult is it to learn how to use the system?
+\\ \hline
+6  & I found the system unnecessarily complex  & I found the system appropriately simple        & How complex is the system? \\ \hline
+7  & I thought the system was easy to use      & I thought the system was hard to use           & How easy or difficult is it to use the system? \\ \hline
+8  & I found the various functions in this system were well integrated & I found the various functions in this system were not well integrated & How integrated are the system’s various functions?
+\\ \hline
+9  & I thought there was too much inconsistency in this system         & I did not think there was too much inconsistency in this system       & How consistent is the system? \\ \hline
+10 & I think that I would like to use this system frequently           & I do not think that I would like to use this system frequently        & How much do you like or dislike the system? \\
+\hline
+\end{tabular}
+\end{table*}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+% \section{Acknowledgments}
 

### 2013-09-10 prep_data.R

@@ -1,21 +1,28 @@
 library(psych)
 library(plyr)
 setwd("~/Dropbox/r-code/SUS/")
-dat=read.csv("medstats_250813.csv")
-dat=read.csv("medstats_280813.csv")
-names(dat)
 
 extractQ = function(x){
-	temp = as.factor(as.character(x[-1]))
-	return(temp)
+    temp = as.factor(as.character(x[-1]))
+    return(temp)
 }
 extractQ2 = function(x){
-	temp = as.character(x[-1])
-	return(temp)
+    temp = as.character(x[-1])
+    return(temp)
 }
 fac2num = function(x){
-	return(as.numeric(as.character(x)))
+    return(as.numeric(as.character(x)))
 }
+
+############################
+###### Med Stats Data  #####
+############################
+
+dat=read.csv("raw_data/medstats_250813.csv")
+dat=read.csv("raw_data/medstats_280813.csv")
+dat=read.csv("raw_data/medstats_06092013.csv")
+names(dat)
+
 
 # Remove duplicates and test responses
 dat = subset(dat, a!="")
@@ -69,7 +76,7 @@ q5=fac2num(factor(q5, labels=c(NA,4,2,0,1,3)))
 sus2=data.frame(q1,q2,q3,q4,q5)
 
 sus2_all=data.frame(sus2, browser, os, res, oe, difficulty, redo, learnsat)[rowSums(is.na(sus2))==0,]
-write.csv(sus2_all, "sus2_data.csv", row.names=F)
+write.csv(sus2_all, "sus2_medstat.csv", row.names=F)
 
 
 
@@ -97,19 +104,19 @@ labAD = function(x, rev=0){
 revsus=subset(dat, select=Q75_2:Q75_15)
 revsus[1,]
 rev_sus=data.frame(
-	p1=labAD(revsus[,1]),
-	p2=labAD(revsus[,2]),
-	p3=labAD(revsus[,3],1),
-	p4=labAD(revsus[,4]),
-	p5=labAD(revsus[,5],1),
-	p6=labAD(revsus[,6]),
-	p7=labAD(revsus[,7],1),
-	p8=labAD(revsus[,8],1),
-	p9=labAD(revsus[,9]),
-	p10=labAD(revsus[,10],1))
+	q1=labAD(revsus[,1]),
+	q2=labAD(revsus[,2]),
+	q3=labAD(revsus[,3],1),
[…]

### 2013-09-10 paper_draft1.Rnw
Started writing on acquiescence bias

@@
 
+In this section we provide strong evidence for acquiescence bias in the SUS by comparing original SUS scores against scores of the the reversed SUS.
+If acquiescence bias exists, we would expect respondents to tend to agree with the statement, independent of whether its tone is positive or negative.
+This would lead to significant difference
 <<eval=FALSE>>=

### 2013-09-11 paper_draft1.Rnw
Updates to evaluation sections

--- 2013-09-05 paper_draft1.Rnw
@@
 library(ggplot2)
+library(grid)
 library(lsr)
 library(psych)
-setwd("~/Dropbox/r-code/SUS/")
-sus2 = read.csv("sus2_data.csv")
-orisus = read.csv("ori_sus_data.csv")
-revsus = read.csv("rev_sus_data.csv")
+setwd("~/Dropbox/r-code/SUS2/")
+m.sus2 = read.csv("sus2_medstat.csv")
+m.orisus = read.csv("ori_sus_medstat.csv")
+m.revsus = read.csv("rev_sus_medstat.csv")
+a.sus2 = read.csv("sus2_algo.csv")
+a.orisus = read.csv("ori_sus_algo.csv")
+a.revsus = read.csv("rev_sus_algo.csv")
 @
@@
 \usepackage{graphics} % for EPS, load graphicx instead
-\usepackage{times}    % comment if you want LaTeX's default font
+%\usepackage{times}    % comment if you want LaTeX's default font
 \usepackage{url}      % llt: nicely formatted URLs
+\usepackage{booktabs}
 
@@
 
-\keywords{}
+\keywords{SUS; questionnaire; surveys}
 
@@
 
-<<>>=
+Some of the most used such questionnaires are QUIS \cite{chin1988}, SUMI \cite{kirakowski1993}, CSUQ \cite{lewis1995}, and SUS \cite{brooke1996}.
+
+The SUS is used extensively to evaluate systems' usability and it is thus important that it provides accurate results.
+Since its introduction in 19XX, there have been critical advancement in the literature on questionnaire and question design regarding the identification of question biases and how to avoid them.
+As a matter of fact, the SUS is vulnerable to several well-known biases as a result of how the items and scale are constructed.
+
+[Insert high-level summary of SUS]
+
+\section{Related Work}
+
+[Maybe some background on the development and testing of the SUS. Describe the original SUS and a few of its incarnations until today.]
+
+\subsection{Survey Biases}
+\subsubsection{Satisficing}
+\subsubsection{Acquiescence}
+\subsubsection{Question order}
+\subsubsection{Social Desirability}
+\subsubsection{Answer Options}
+\subsubsection{Hypotheticals}
+\subsubsection{Leading Information}
+
+\section{SUS Evaluation}
+
+\subsection{Heuristic review}
+In this section, we inspect the SUS, one item at a time, for deficiencies in its question design. The item numbers that are refered to correspond to those in Table \ref{tab:items}.
+[Go through each question and describe biases, with references to articles that described them]
+
+\subsection{Experiment setup}
+Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey.
+The survey received 1746 responses.
+At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, their satisfaction with the amount they learnt, and the difficulty of the course.
+Respondents were then randomly assigned to one of three weighted groups: 25\% were presented with the original SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received the SUS 2.0 proposed in this paper (n=869).
+The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
+The rest of the survey was the same for all respondents and contained typical course assessment questions.
+
+\subsection{Psychometric Properties of the SUS}
+
+Table \ref{tab:dist} provides basic statistics that describe the distribution of the SUS scores.
+
+\subsubsection{Internal consistency}
+For the 439 responses to the original SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence.
+Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor}, an alpha of 0.86 reflects a high degree of interrelatedness.
+<<eval=FALSE>>=
+alpha(m.orisus[,1:10])$total
+for(i in 1:10){
+    print(cor.test(rowMeans(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
+}
+@
+
+\subsubsection{Factor analysis}
+Lewis \& Sauro \cite{lewis2009factor} investigated the factor structure of the SUS and found two moderately correlated factors: Usability (items 2-4, 6-10) and Learnability (items 1, 5).
+A factor analysis of our 439 responses suggest that a single-factor solution is viable, even though two factors have eigenvalues greater than one.
+A scree plot (Figure \ref{fig:scree}) illustrates this result.
+
+<<eval=FALSE>>=
+fa.parallel(m.orisus[,1:10])
+scree(m.orisus[,1:10])
+fa(m.orisus[,1:10],1)$e.values
+@
+
+\subsection{Acquiescence Bias in SUS}
+
+A comparison of original SUS scores with scores from the reversed SUS provides strong evidence for acquiescence bias in the SUS.
+Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
+However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant differenece between the original SUS average and reversed SUS reverse-coded average.
+
+Table \ref{tab:acqui} provides means, standard deviations, and significance levels from Welch two-sample tests (which does not assume equal sample variances like the t-test).
+Given the much larger sample size of responses on system A than B, we observe more significant differences for comparisons for system A than B.
+Most notably, highly significant acquiescence bias at $p<0.05$ was found for 6 out of 10 items and the overall SUS score for system A.
+This result clearly shows that the SUS is significantly biased.
+
+\begin{table}[h]
+\small
+\centering
+\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
+\label{tab:acqui}
+\begin{tabular}{rccccr}
+\toprule
+ & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+\# & M & SD & M & SD & p value \\
+\midrule
+1 & 8.52 & 2.12 & 6.69 & 2.98 & \textless0.001\\
+2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
+3 & 8.58 & 1.79 & 8.90 & 1.83 & 0.008\\
+4 & 8.19 & 2.36 & 8.66 & 1.68 & \textless0.001\\
+5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.67\\
+6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.49\\
+7 & 8.61 & 1.68 & 8.90 & 1.79 & 0.012\\
+8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.13\\
+9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
+10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.056\\
+\midrule
+ & 84.46 & 12.91 & 82.33 & 11.39 & 0.010\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\section{SUS 2.0 Proposal}
+
+[Describe the proposal and refer to table with the updated questions. Mention that it was a non-goal to reduce the number of questions, etc.]
+
+[We gotta talk about how this impacts the SUS score calculations, comparison to past scores, etc.]
+
+\subsection{Psychometric Properties of the SUS 2.0}
+
+Distribution in Table \ref{tab:dist}
+
+\subsubsection{Internal consistency}
+Coefficient $\alpha$ is 0.77
+<<eval=FALSE>>=
+alpha(m.sus2[,1:5])$total
+@
+
+\subsubsection{Factor analysis}
+Figure \ref{fig:scree}
+<<eval=FALSE>>=
+fa(m.sus2[,1:5],3)$loadings
+fa(m.sus2[,1:5],2)$loadings
+fa(m.sus2[,1:5],1)$loadings
+@
+
+\subsection{Sensitivity of the SUS 2.0}
+[Introduce second data set]
+The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design.
+Based on a design heuristic evaluation of the two systems, it was determined that B had more usability problems than A.
+[Need to talk about how this was evaluated.
+E.g.
+\url{http://en.wikipedia.org/wiki/Heuristic_evaluation}]
+
+<<eval=FALSE>>=
+# testItems=function(i, oa, ra, ob, rb){
+#     print(paste(i, round(t.test(oa,ra)$p.value,3), round(t.test(ob,rb)$p.value,3), sep=", "))
+#     print(paste(i, paste(round(mean(oa),2), " (", round(sd(oa),2), ")", sep=""), paste(round(mean(ra),2), " (", round(sd(ra),2), ")", sep=""),
+#         paste(round(mean(ob),2), " (", round(sd(ob),2), ")", sep=""), paste(round(mean(rb),2), " (", round(sd(rb),2), ")", "\\", sep=""), sep=" & "))
+# }
+# for(i in 1:10){
+#     testItems(i, m.orisus[,i], m.revsus[,i], a.orisus[,i], a.revsus[,i])
+# }
+# testItems("all", rowMeans(m.orisus[,1:10]), rowMeans(m.revsus[,1:10]), rowMeans(a.orisus[,1:10]), rowMeans(a.revsus[,1:10]))
+
+testItems=function(i, oa, ra){
+    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
+}
+for(i in 1:10){
+    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
+}
+testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
+
+@
+
+<<eval=FALSE>>=
+alpha(m.orisus[,1:10])
+fa.diagram(fa(ori_sus,1))
+fa.diagram(fa(ori_sus,2))
+
+cor.test(oe,rowSums(ori_sus))
+cor.test(difficulty,rowSums(ori_sus))
+cor.test(redo,rowSums(ori_sus))
+cor.test(learnsat,rowSums(ori_sus))
+
+# Original SUS
+# I needed to learn a lot of things before I could get going with this system
+# I think that I would need the support of a technical person to be able to use this system
+# I felt very confident using the system
+# I found the system very cumbersome to use
+# I would imagine that most people would learn to use this system very quickly
+# I found the system unnecessarily complex
+# I thought the system was easy to use
+# I found the various functions in this system were well integrated
+# I thought there was too much inconsistency in this system
+# I think that I would like to use this system frequently
+@
+
+<<eval=FALSE>>=
 # SUS 2.0
@@
 cor.test(learnsat,rowSums(ori_sus))
+@
 
-t.test(ori_sus$p1, rev_sus$p1)
-t.test(ori_sus$p2, rev_sus$p2)
-t.test(ori_sus$p3, rev_sus$p3)
-t.test(ori_sus$p4, rev_sus$p4)
-t.test(ori_sus$p5, rev_sus$p5) #not
-t.test(ori_sus$p6, rev_sus$p6) #not
-t.test(ori_sus$p7, rev_sus$p7)
-t.test(ori_sus$p8, rev_sus$p8) #not
-t.test(ori_sus$p9, rev_sus$p9)
-t.test(ori_sus$p10, rev_sus$p10)
+\begin{table*}
+\small
+\centering
+\caption{Items from the original and reversed SUS}
+\label{tab:items}
+\begin{tabular}{| r | p{8cm} | p{8cm} |}
+\hline
+\textbf{\#}  & \textbf{Original SUS$^\ast$} & \textbf{Reversed SUS$^\ast$}\\
+\hline
+1  & I needed to learn a lot of things before I could get going with this system   & I needed to learn very few things before I could get going with this system \\ \hline
+2  & I think that I would need the support of a technical person to be able to use this system & I think that I would not need any support of a technical person to be able to use this system \\ \hline
+3  & I felt very confident using the system    & I did not feel very confident using the system\\ \hline
+4  & I found the system very cumbersome to use & I found the system very manageable to use\\ \hline
+5  & I would imagine that most people would learn to use this system very quickly  & I would imagine that most people would learn to use this system very slowly\\ \hline
+6  & I found the system unnecessarily complex  & I found the system appropriately simple\\ \hline
+7  & I thought the system was easy to use      & I thought the system was hard to use\\ \hline
+8  & I found the various functions in this system were well integrated & I found the various functions in this system were not well integrated\\ \hline
+9  & I thought there was too much inconsistency in this system         & I did not think there was too much inconsistency in this system \\ \hline
+10 & I think that I would like to use this system frequently           & I do not think that I would like to use this system frequently \\
+\hline
+\end{tabular}
+\smallskip
+\scriptsize{$^\ast$Items were presented in a matrix with a 5-point Likert scale: Strongly disagree (1), (2), (3), (4), Strongly agree (5)}
+\end{table*}
 
-t.test(rowMeans(ori_sus), rowMeans(rev_sus))
+\begin{table*}
+\small
+\centering
+\caption{Items from the proposed SUS 2.0 with answer scales}
+\label{tab:sus2items}
+\begin{tabular}{| r | p{7cm} | p{9cm} |}
+\hline
+\textbf{\#}  & \textbf{Questions} & \textbf{Answer Scales} \\
+\hline
+1$^\star$  & How much more is there to learn about the system? & Nothing at all, A little, A moderate amount, A lot, A great deal \\ \hline
+2  & How likely are you to need the support of a technical person to be able to use the system?
+& \{Extremely, Very, Somewhat\} unlikely, Neither likely nor unlikely, \{Somewhat, Very, Extremely\} likely \\ \hline
+3$^\star$  & How confident are you using the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} confident \\ \hline
+4  & How cumbersome is it to use the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome \\ \hline
+5$^\star$  & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
+6  & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex \\ \hline
+7$^\star$  & How easy or difficult is it to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
+8  & How integrated are the system’s various functions? & \{Not at all, Slightly, Moderately, Very, Extremely\} integrated \\ \hline
+9$^\star$  & How consistent is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} consistent \\ \hline
+10 & How much do you like or dislike the system? & \{Extremely, Moderately, Slightly\} dislike, Neither like nor dislike, \{Slightly, Moderately, Extremely\} like\\
+\hline
+\end{tabular}
+\end{table*}
 
+<<eval=FALSE>>=
+convSUS = function(x) {
+    return(rowSums((x+1)*2))
+}
+convSUS2 = function(x) {
+    return(10*(as.matrix(m.sus2[,1:5]+1) %*% (c(5,5,7,7,5)/20))[,1])
+}
+length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
+length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10]))
+length(convSUS2(m.sus2[,1:5])); range(convSUS2(m.sus2[,1:5])); mean(convSUS2(m.sus2[,1:5])); sd(convSUS2(m.sus2[,1:5])); median(convSUS2(m.sus2[,1:5])); IQR(convSUS2(m.sus2[,1:5]))
 @
+
+\begin{table}
+\small
+\centering
+\caption{Statistical information on the SUS distributions (SUS 2.0 scores scaled to 0-100 for comparison)}
+\label{tab:dist}
+\begin{tabular}{lccccccc}
+\toprule
+SUS & N & Min & Max & Mean & SD & Median & IQR\\
+\midrule
+Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
+Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
+2.0        & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
+scree=data.frame(
+    n=c(1:10,1:5), 
+    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
+    sus=rep(c("Original SUS", "SUS 2.0"),c(10,5)))
+ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.7,.75))
+@
+
+\section{Discussion}
+
+\section{Conclusion}
 

### 2013-09-12 paper_draft1.Rnw
Updated sensitivity part

@@
 library(grid)
+library(boot)
 library(lsr)
@@
 a.revsus = read.csv("rev_sus_algo.csv")
+SE=function(x){
+    return(sqrt(var(x)/length(x)))
+}
+convSUS = function(x) {
+    return(rowSums((x+1)*2))
+}
+convSUS2 = function(x) {
+    return(10*(as.matrix(x+1) %*% (c(5,5,7,7,5)/20))[,1])
+}
+convSUS2s = function(x) {
+    return(10*(as.matrix(x+1) %*% (c(5,7,7,5)/25))[,1])
+}
 @
@@
 \begin{abstract}
-
+The System Usability Scale (SUS) is probably the most widely employed measure of usability.
+Numerous publications have been devoted to assessing its psychometric properties and using it as a gold standard in the development of alternative metrics.
+From a questionnaire design perspective, however, the SUS has striking deficiencies which lead to biased measures of usability.
+First, we review the literature on survey biases, inspect each SUS item for issues and show that the SUS is vulnerable to significant acquiescence bias using a survey experiment.
+We then propose the SUS 2.0, an updated version of the SUS, which conforms to insights from questionnaire design research, and present its favorable psychometric properties.
 \end{abstract}
@@
 
-Some of the most used such questionnaires are QUIS \cite{chin1988}, SUMI \cite{kirakowski1993}, CSUQ \cite{lewis1995}, and SUS \cite{brooke1996}.
+With the increased focus on developing products that have a high level of usability, it has become important to quantify the perceived usability of a product or system.
+Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
+Commonly used such questionnaires include the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the Software Usability Measurement Inventory (SUMI) \cite{kirakowski1993}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among many others.
+Out of all usability questionnaires developed, SUS has received the highest level of adoption in both industry and academia, with hundreds of references in various publications alone.
 
-The SUS is used extensively to evaluate systems' usability and it is thus important that it provides accurate results.
-Since its introduction in 19XX, there have been critical advancement in the literature on questionnaire and question design regarding the identification of question biases and how to avoid them.
-As a matter of fact, the SUS is vulnerable to several well-known biases as a result of how the items and scale are constructed.
+The SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation in the UK.
+It was used as a “quick and dirty” scale to administer after usability studies on electronic office systems, such as the VT100, a text-based terminal system.
+SUS measures attitudes and perceptions regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11), yielding a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
+To measure the system’s usability on these dimensions, SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate on a five-point Likert scale.
+The Likert scale, established by Rensis Likert in 1932, allows questionnaire respondents to specify their level of agreement or disagreement on a symmetric agreement scale for a series of statements \cite{likert1932}.
+SUS’ Likert scale ranges from “Strongly disagree” to “Strongly agree”, with only its endpoints labeled, while additionally all five scale items are numbered from 1 to 5.
+As already noted in Brooke’s initial work \cite{brooke1996}, the phrasing of the statement heavily influences the respondent’s level of agreement or disagreement for it.
+During the development of the SUS, statements that received the most extreme responses were selected.
+Finally, when analyzing the SUS responses, the individual responses are added up using a particular scheme.
+To ensure that this summation is possible, all items of the SUS need to be evaluated by the respondent.
 
-[Insert high-level summary of SUS]
+Over the years, SUS has been used across a variety of different systems, including hardware, software, websites, and applications.
+However, since the time of its development in 1986, research regarding questionnaire design has advanced significantly, with several insights that now challenge some of the foundations of the SUS.
+The remainder of this note will explain relevant advances in questionnaire design research, evaluate the original SUS in the context of those, propose an updated version to conform with these insights, and finally offer a comparison between the original and the updated SUS.
+Note that our intention is not to reduce the number of statements asked about in the SUS, contrary to recent work that attempts to create a shorter usability measure to save time \cite{lewis2013,finstad2010usability}.
+Our goal is to merely update the original questionnaire.
 
@@
 
+As the psychometric propoerties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics is sufficient for this paper.
 Table \ref{tab:dist} provides basic statistics that describe the distribution of the SUS scores.
@@
 For the 439 responses to the original SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence.
-Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor}, an alpha of 0.86 reflects a high degree of interrelatedness.
+Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an alpha of 0.86 reflects a high degree of interrelatedness.
 <<eval=FALSE>>=
@@
 \subsubsection{Factor analysis}
-Lewis \& Sauro \cite{lewis2009factor} investigated the factor structure of the SUS and found two moderately correlated factors: Usability (items 2-4, 6-10) and Learnability (items 1, 5).
+%Lewis \& Sauro \cite{lewis2009factor} investigated the factor structure of the SUS and found two moderately correlated factors: Usability (items 2-4, 6-10) and Learnability (items 1, 5). 
 A factor analysis of our 439 responses suggest that a single-factor solution is viable, even though two factors have eigenvalues greater than one.
@@
 \subsection{Acquiescence Bias in SUS}
-
 A comparison of original SUS scores with scores from the reversed SUS provides strong evidence for acquiescence bias in the SUS.
@@
 \end{table}
+<<eval=FALSE>>=
+# testItems=function(i, oa, ra, ob, rb){
+#     print(paste(i, round(t.test(oa,ra)$p.value,3), round(t.test(ob,rb)$p.value,3), sep=", "))
+#     print(paste(i, paste(round(mean(oa),2), " (", round(sd(oa),2), ")", sep=""), paste(round(mean(ra),2), " (", round(sd(ra),2), ")", sep=""),
+#         paste(round(mean(ob),2), " (", round(sd(ob),2), ")", sep=""), paste(round(mean(rb),2), " (", round(sd(rb),2), ")", "\\", sep=""), sep=" & "))
+# }
+# for(i in 1:10){
+#     testItems(i, m.orisus[,i], m.revsus[,i], a.orisus[,i], a.revsus[,i])
+# }
+# testItems("all", rowMeans(m.orisus[,1:10]), rowMeans(m.revsus[,1:10]), rowMeans(a.orisus[,1:10]), rowMeans(a.revsus[,1:10]))
+
+testItems=function(i, oa, ra){
+    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
+}
+for(i in 1:10){
+    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
+}
+testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
+
+@
 
@@
 
+Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for two online course interfaces where one interface has an objectively higher usability level than the other.
+As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.
+While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity.
+A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0)
 <<eval=FALSE>>=
-# testItems=function(i, oa, ra, ob, rb){
-#     print(paste(i, round(t.test(oa,ra)$p.value,3), round(t.test(ob,rb)$p.value,3), sep=", "))
-#     print(paste(i, paste(round(mean(oa),2), " (", round(sd(oa),2), ")", sep=""), paste(round(mean(ra),2), " (", round(sd(ra),2), ")", sep=""),
-#         paste(round(mean(ob),2), " (", round(sd(ob),2), ")", sep=""), paste(round(mean(rb),2), " (", round(sd(rb),2), ")", "\\", sep=""), sep=" & "))
-# }
-# for(i in 1:10){
-#     testItems(i, m.orisus[,i], m.revsus[,i], a.orisus[,i], a.revsus[,i])
-# }
-# testItems("all", rowMeans(m.orisus[,1:10]), rowMeans(m.revsus[,1:10]), rowMeans(a.orisus[,1:10]), rowMeans(a.revsus[,1:10]))
+wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]))
+wilcox.test(convSUS(m.sus2[,2:5]), convSUS(a.sus2[,2:5]))
+@
 
-testItems=function(i, oa, ra){
-    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
-}
-for(i in 1:10){
-    testItems(i, (2*m.orisus[,i]+2), (2*m.revsus[,i]+2))
-}
-testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
+<<sens, fig.height=3.5, fig.width=4, fig.cap='Evaluation of scale sensitivity for original SUS and SUS 2.0 showing that the SUS 2.0 has high enough sensitivity to distinguish between a high and low usability system'>>=
+subMean = function(data, indices) {mean(data[indices])}
+m.oriBoot = boot(data=convSUS(m.orisus[,1:10]), statistic=subMean, R=10000)
+a.oriBoot = boot(data=convSUS(a.orisus[,1:10]), statistic=subMean, R=10000)
+m.sus2Boot = boot(data=rowMeans(m.sus2[,2:5]), statistic=subMean, R=10000)
+a.sus2Boot = boot(data=rowMeans(a.sus2[,2:5]), statistic=subMean, R=10000)
 
+# boot.ci(m.oriBoot, type="bca") # (83.21, 85.64 )  
+# boot.ci(a.oriBoot, type="bca") # (84.90, 89.29 ) 
+# boot.ci(m.sus2Boot, type="bca") # ( 3.404,  3.494 )
+# boot.ci(a.sus2Boot, type="bca") # ( 3.543,  3.702 )
+
+sensDat=data.frame(
+    sus=c("\nOriginal SUS", "\nOriginal SUS", "\nShort SUS 2.0", "\nShort SUS 2.0"),
+    system=c("Low Usability System","High Usability System  ","Low Usability System","High Usability System  "),
+    score=c(m.oriBoot$t0, a.oriBoot$t0, m.sus2Boot$t0, a.sus2Boot$t0),
+    lwr=c(83.21, 84.90, 3.404, 3.543),
+    upr=c(85.64, 89.29, 3.494, 3.702))
+
+ggplot(sensDat, aes(x=score, y=sus, xmin=lwr, xmax=upr, linetype=system)) +
+    geom_point(size=2) +
+    geom_errorbarh(height=.5) +
+    theme_bw() +
+    facet_wrap(~sus, scales="free", nrow=2) +
+    theme(legend.position="bottom", axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
+    labs(x="Scores with 95% C.I.", y="", linetype="")
 @
@@
 
-# Original SUS
-# I needed to learn a lot of things before I could get going with this system
-# I think that I would need the support of a technical person to be able to use this system
-# I felt very confident using the system
-# I found the system very cumbersome to use
-# I would imagine that most people would learn to use this system very quickly
-# I found the system unnecessarily complex
-# I thought the system was easy to use
-# I found the various functions in this system were well integrated
-# I thought there was too much inconsistency in this system
-# I think that I would like to use this system frequently
 @
@@
 <<eval=FALSE>>=
-# SUS 2.0
-# q1 How easy or difficult was it to learn how to use the system?
-# q2 How much more is there to learn about the system?
-# q3 How confident were you using the system?
-# q4 How easy or difficult is it to use the system?
-# q5 How consistent is the system?
-
 # Criterion 
@@
 
-fa.parallel(sus2)
-
-# REV SUS
-# I needed to learn very few things before I could get going with this system
-# I think that I would not need any support of a technical person to be able to use this system
-# I did not feel very confident using the system
-# I found the system very manageable to use
-# I would imagine that most people would learn to use this system very slowly
-# I found the system appropriately simple
-# I thought the system was hard to use
-# I found the various functions in this system were not well integrated
-# I did not think there was too much inconsistency in this system
-# I do not think that I would like to use this system frequently
-
-alpha(rev_sus)
-fa.diagram(fa(rev_sus,1))
-fa.diagram(fa(rev_sus,2))
-
-# Original SUS
-# I needed to learn a lot of things before I could get going with this system
-# I think that I would need the support of a technical person to be able to use this system
-# I felt very confident using the system
-# I found the system very cumbersome to use
-# I would imagine that most people would learn to use this system very quickly
-# I found the system unnecessarily complex
-# I thought the system was easy to use
-# I found the various functions in this system were well integrated
-# I thought there was too much inconsistency in this system
-# I think that I would like to use this system frequently
-
-alpha(ori_sus)
-fa.diagram(fa(ori_sus,1))
-fa.diagram(fa(ori_sus,2))
-
 cor.test(oe,rowSums(ori_sus))
@@
 \label{tab:items}
-\begin{tabular}{| r | p{8cm} | p{8cm} |}
+\begin{tabular}{| l | p{8cm} | p{8cm} |}
 \hline
@@
 \label{tab:sus2items}
-\begin{tabular}{| r | p{7cm} | p{9cm} |}
+\begin{tabular}{| l | p{7cm} | p{9cm} |}
 \hline
@@
 \end{tabular}
+\smallskip
+\scriptsize{$^\ast$Items included in the Short SUS 2.0}
 \end{table*}
@@
 <<eval=FALSE>>=
-convSUS = function(x) {
-    return(rowSums((x+1)*2))
-}
-convSUS2 = function(x) {
-    return(10*(as.matrix(m.sus2[,1:5]+1) %*% (c(5,5,7,7,5)/20))[,1])
-}
 length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
@@
 Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
-2.0        & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
+Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
 \bottomrule
@@
     eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
-    sus=rep(c("Original SUS", "SUS 2.0"),c(10,5)))
-ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.7,.75))
+    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
+ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
 @

### 2013-09-13 paper_draft1.Rnw
added more details to analysis

@@
 a.revsus = read.csv("rev_sus_algo.csv")
-SE=function(x){
+SE = function(x){
     return(sqrt(var(x)/length(x)))
@@
 The System Usability Scale (SUS) is probably the most widely employed measure of usability.
-Numerous publications have been devoted to assessing its psychometric properties and using it as a gold standard in the development of alternative metrics.
+Numerous studies have assessed its psychometric properties and used it as a `gold standard' in the development of alternative scales.
 From a questionnaire design perspective, however, the SUS has striking deficiencies which lead to biased measures of usability.
@@
 We then propose the SUS 2.0, an updated version of the SUS, which conforms to insights from questionnaire design research, and present its favorable psychometric properties.
+We present evidence from a second study that suggests that the SUS 2.0 is a more sensitive usability measure than the original SUS.
 \end{abstract}
@@
 
+\begin{table}
+\small
+\centering
+\caption{Statistical information on the SUS distributions (short SUS 2.0 scores scaled to 0-100 for comparison)}
+\label{tab:dist}
+\begin{tabular}{lcccccccc}
+\toprule
+SUS & N & Min & Max & Mean & SD & Median & IQR\\
+\midrule
+Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
+Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
+Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsubsection{Concurrent Validity}
+We expect the original SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs.
+We find that the original SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.27, $t(437)$=6, $p\textless$0.001).
+However, there we find no significant association between the SUS and the difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4).
+<<eval=FALSE>>=
+cor.test(m.orisus$oe,rowSums(m.orisus[,1:10]))
+cor.test(m.orisus$redo,rowSums(m.orisus[,1:10]))
+cor.test(m.orisus$learnsat,rowSums(m.orisus[,1:10]))
+cor.test(m.orisus$difficulty,rowSums(m.orisus[,1:10]))
+@
+
 \subsubsection{Internal consistency}
@@
 alpha(m.orisus[,1:10])$total
+omega(m.orisus[,1:10], nfactors=2)
+iclust(m.orisus[,1:10],nclusters=2)
 for(i in 1:10){
@@
 \subsubsection{Factor analysis}
-%Lewis \& Sauro \cite{lewis2009factor} investigated the factor structure of the SUS and found two moderately correlated factors: Usability (items 2-4, 6-10) and Learnability (items 1, 5). 
-A factor analysis of our 439 responses suggest that a single-factor solution is viable, even though two factors have eigenvalues greater than one.
-A scree plot (Figure \ref{fig:scree}) illustrates this result.
-
+A factor analysis of the 439 responses to the original SUS suggests that the scale has a two-factor structure.
+A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one.
+This is consistent with previous work on the SUS factor structure \cite{lewis2009factor}.
 <<eval=FALSE>>=
@@
 scree(m.orisus[,1:10])
-fa(m.orisus[,1:10],1)$e.values
+fa(m.orisus[,1:10],2)$e.values
+fa.diagram(fa(m.orisus[,1:10],2))
+@
+<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
+scree=data.frame(
+    n=c(1:10,1:5), 
+    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
+    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
+ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
 @
@@
 <<eval=FALSE>>=
-# testItems=function(i, oa, ra, ob, rb){
-#     print(paste(i, round(t.test(oa,ra)$p.value,3), round(t.test(ob,rb)$p.value,3), sep=", "))
-#     print(paste(i, paste(round(mean(oa),2), " (", round(sd(oa),2), ")", sep=""), paste(round(mean(ra),2), " (", round(sd(ra),2), ")", sep=""),
-#         paste(round(mean(ob),2), " (", round(sd(ob),2), ")", sep=""), paste(round(mean(rb),2), " (", round(sd(rb),2), ")", "\\", sep=""), sep=" & "))
-# }
-# for(i in 1:10){
-#     testItems(i, m.orisus[,i], m.revsus[,i], a.orisus[,i], a.revsus[,i])
-# }
-# testItems("all", rowMeans(m.orisus[,1:10]), rowMeans(m.revsus[,1:10]), rowMeans(a.orisus[,1:10]), rowMeans(a.revsus[,1:10]))
-
 testItems=function(i, oa, ra){
@@
 testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
-
 @
@@
 
-Distribution in Table \ref{tab:dist}
+Table \ref{tab:dist} provides basic statistics on the distribution of short SUS 2.0 scores scaled to the same 0-100 score range as the original SUS to facilitate a comparison.
+Most notably, the SUS 2.0 scores range from about 25 to 85 for a system with good usability, while the original and reversed SUS suffer from ceiling effects which artificially constrain the distribution of scores.
+
+\subsubsection{Concurrent Validity}
+We start to establish the SUS 2.0's concurrent validity by investigating its association with related and unrelated constructs.
+We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.22, $t(867)$=7, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.25, $t(867)$=8, $p\textless$0.001); however, the short SUS 2.0 is only marignally associated with the difficulty of the course ($r$=0.08, $t(867)$=2.5, p=0.01).
+<<eval=FALSE>>=
+cor.test(m.sus2$oe,rowSums(m.sus2[,1:5]))
+cor.test(m.sus2$learnsat,rowSums(m.sus2[,1:5]))
+cor.test(m.sus2$redo,rowSums(m.sus2[,1:5]))
+cor.test(m.sus2$difficulty,rowSums(m.sus2[,1:5]))
+@
 
 \subsubsection{Internal consistency}
-Coefficient $\alpha$ is 0.77
+Cronbach's $\alpha$ for the short SUS 2.0 is 0.77 based on 869 responses and individual item correlations with the total score vary between 0.47 and 0.86 with 95\% confidence.
+Coefficient alpha is smaller for the short SUS 2.0, but given that coefficient alpha increases with the number of items, the ten-item SUS 2.0 will have higher reliability thant the five-item short SUS 2.0.
 <<eval=FALSE>>=
 alpha(m.sus2[,1:5])$total
+omega(m.orisus[,1:10],nfactors=2)
+for(i in 1:5){
+    print(cor.test(rowMeans(m.sus2[,1:5]), m.sus2[,i])$conf[1:2])
+}
 @
@@
 \subsubsection{Factor analysis}
-Figure \ref{fig:scree}
+A factor analysis of 869 responses to the short SUS 2.0 suggests that a single-factor structure is most appropriate for the measure.
+As shown in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point.
 <<eval=FALSE>>=
@@
 \subsection{Sensitivity of the SUS 2.0}
-[Introduce second data set]
+A good usability scale should exhibit a high level of sensitivity to reflect subtle differences in usability.
+We conducted a second survey study to investigate how the SUS 2.0 compares to the original SUS in terms of sensitivity.
 The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design.
@@
 While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity.
-A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0)
+A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0).
 <<eval=FALSE>>=
@@
 <<eval=FALSE>>=
-alpha(m.orisus[,1:10])
-fa.diagram(fa(ori_sus,1))
-fa.diagram(fa(ori_sus,2))
-
 cor.test(oe,rowSums(ori_sus))
@@
 
-\begin{table}
-\small
-\centering
-\caption{Statistical information on the SUS distributions (SUS 2.0 scores scaled to 0-100 for comparison)}
-\label{tab:dist}
-\begin{tabular}{lccccccc}
-\toprule
-SUS & N & Min & Max & Mean & SD & Median & IQR\\
-\midrule
-Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
-Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
-Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
-\bottomrule
-\end{tabular}
-\end{table}
-
-<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
-scree=data.frame(
-    n=c(1:10,1:5), 
-    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
-    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
-ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
-@
-
 \section{Discussion}

### 2013-09-13 paper_draft1.Rnw
added info on sus2.0 development

@@
 
-\begin{table}
+\begin{table}[b]
 \small
@@
 
-[Describe the proposal and refer to table with the updated questions. Mention that it was a non-goal to reduce the number of questions, etc.]
+Following the heuristic evaluation of the original SUS items, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias.
+The first step was to change the questionnaire items from being statements to questions in an effort to reduce acquiescence bias.
+Moreover, statements that were phrased as hypotheticals, such as item 5 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying construct.
+
+The second step was to change the scale from an agree-disagree Likert scale to scales that reflect the relevant constructs, such as confidence, learnability, or complexity.
+Following question design research, unipolar scales were presented as 5-point scales, while bipolar scales were presented as 7-point scales.
+
+Our goal was to update the SUS items for future use given that the SUS is clearly the most established usability scale.
+These updated items with corresponding answer scales are presented in Table \ref{tab:sus2items}.
+For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
+The short SUS 2.0 consists of 5 items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (leanability, confidence, ease of use, consistency).
 
@@
 A good usability scale should exhibit a high level of sensitivity to reflect subtle differences in usability.
-We conducted a second survey study to investigate how the SUS 2.0 compares to the original SUS in terms of sensitivity.
-The two online courses were offered on two distinct online platforms that shared the same core features but differed considerably in design.
-Based on a design heuristic evaluation of the two systems, it was determined that B had more usability problems than A.
-[Need to talk about how this was evaluated.
-E.g.
-\url{http://en.wikipedia.org/wiki/Heuristic_evaluation}]
+We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how the SUS 2.0 compares to the original SUS in terms of sensitivity.
+
+The two online courses that are compared were offered on two distinct online platforms that shared the same core features but differed considerably in design.
+Based on Nielsen's heuristic evaluation of the two systems, it was determined that one system had more usability problems than the other.
+The following criteria were used [go through these and say which ones were different/better etc.]
+
+Visibility of system status:
+%The system should always keep users informed about what is going on, through appropriate feedback within reasonable time.
+
+Match between system and the real world:
+%The system should speak the user's language, with words, phrases and concepts familiar to the user, rather than system-oriented terms.
+Follow real-world conventions, making information appear in a natural and logical order.
+
+User control and freedom:
+%Users often choose system functions by mistake and will need a clearly marked "emergency exit" to leave the unwanted state without having to go through an extended dialogue. Support undo and redo.
+
+Consistency and standards:
+%Users should not have to wonder whether different words, situations, or actions mean the same thing. Follow platform conventions.
+
+Error prevention:
+%Even better than good error messages is a careful design which prevents a problem from occurring in the first place.
+Either eliminate error-prone conditions or check for them and present users with a confirmation option before they commit to the action.
+
+Recognition rather than recall:
+%Minimize the user's memory load by making objects, actions, and options visible.
+The user should not have to remember information from one part of the dialogue to another.
+Instructions for use of the system should be visible or easily retrievable whenever appropriate.
+
+Flexibility and efficiency of use:
+%Accelerators—unseen by the novice user—may often speed up the interaction for the expert user such that the system can cater to both inexperienced and experienced users.
+Allow users to tailor frequent actions.
+
+Aesthetic and minimalist design:
+%Dialogues should not contain information which is irrelevant or rarely needed.
+Every extra unit of information in a dialogue competes with the relevant units of information and diminishes their relative visibility.
+
+Help users recognize, diagnose, and recover from errors:
+%Error messages should be expressed in plain language (no codes), precisely indicate the problem, and constructively suggest a solution.
+
+Help and documentation:
+%Even though it is better if the system can be used without documentation, it may be necessary to provide help and documentation.
+Any such information should be easy to search, focused on the user's task, list concrete steps to be carried out, and not be too large.
 
@@
 <<eval=FALSE>>=
-cor.test(oe,rowSums(ori_sus))
-cor.test(difficulty,rowSums(ori_sus))
-cor.test(redo,rowSums(ori_sus))
-cor.test(learnsat,rowSums(ori_sus))
-
+length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
+length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10]))
+length(convSUS2(m.sus2[,1:5])); range(convSUS2(m.sus2[,1:5])); mean(convSUS2(m.sus2[,1:5])); sd(convSUS2(m.sus2[,1:5])); median(convSUS2(m.sus2[,1:5])); IQR(convSUS2(m.sus2[,1:5]))
 @
 
-<<eval=FALSE>>=
-# Criterion 
-cor.test(oe,rowSums(sus2))
-cor.test(difficulty,rowSums(sus2))
-cor.test(redo,rowSums(sus2))
-cor.test(learnsat,rowSums(sus2))
+\section{Discussion}
 
-# Internal Consistency
-cor.test(rowSums(sus2), q1)
-cor.test(rowSums(sus2), q2)
-cor.test(rowSums(sus2), q3)
-cor.test(rowSums(sus2), q4)
-cor.test(rowSums(sus2), q5)
+\section{Conclusion}
 
-# Cronbach's Alpha
-alpha(sus2)
-
-# Split Half Reliability
-rhh=cor(rowSums(cbind(q1,q3,q5), na.rm=T), rowSums(cbind(q2,q4), na.rm=T))
-(2*rhh)/(1+rhh)
-
-# Factor Analysis
-plot(princomp(~q1+q2+q3+q4+q5, data=sus2, scores=T), type="lines")
-
-eigen(cov(sus2, use="pair"))
-eigen.loadings(eigen(cov(sus2, use="pair")))
-princomp(covmat=cov(sus2, use="pair"))$loadings
-plot(princomp(covmat=cov(sus2, use="pair")), type="lines")
-
-cor.test(oe,rowSums(ori_sus))
-cor.test(difficulty,rowSums(ori_sus))
-cor.test(redo,rowSums(ori_sus))
-cor.test(learnsat,rowSums(ori_sus))
-@
+\section{Acknowledgments}
+We are grateful to the instructors of the two Stanford courses who allowed us conduct survey experiments in their post-course surveys.
 
@@
 
-<<eval=FALSE>>=
-length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
-length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10]))
-length(convSUS2(m.sus2[,1:5])); range(convSUS2(m.sus2[,1:5])); mean(convSUS2(m.sus2[,1:5])); sd(convSUS2(m.sus2[,1:5])); median(convSUS2(m.sus2[,1:5])); IQR(convSUS2(m.sus2[,1:5]))
-@
-
-\section{Discussion}
-
-\section{Conclusion}
-
-\section{Acknowledgments}
-
 \bibliographystyle{acm-sigchi}

### 2013-09-14 paper_draft1.Rnw
replaced t-tests with WM tests, wrote discussion and conclusion

@@
 
-Table \ref{tab:acqui} provides means, standard deviations, and significance levels from Welch two-sample tests (which does not assume equal sample variances like the t-test).
-Given the much larger sample size of responses on system A than B, we observe more significant differences for comparisons for system A than B.
-Most notably, highly significant acquiescence bias at $p<0.05$ was found for 6 out of 10 items and the overall SUS score for system A.
-This result clearly shows that the SUS is significantly biased.
+Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (this non-parametric alternative of the t-test is used as scores are not normally distributed).
+We find highly significant differences at 99\% significance in all but two items and the overall SUS score.
+This is very strong evidence for the claim that the original SUS induces acquiescence bias.
 
@@
 2 & 9.09 & 1.85 & 8.15 & 2.47 & \textless0.001\\
-3 & 8.58 & 1.79 & 8.90 & 1.83 & 0.008\\
-4 & 8.19 & 2.36 & 8.66 & 1.68 & \textless0.001\\
-5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.67\\
-6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.49\\
-7 & 8.61 & 1.68 & 8.90 & 1.79 & 0.012\\
-8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.13\\
+3 & 8.58 & 1.79 & 8.90 & 1.83 & \textless0.001\\
+4 & 8.19 & 2.36 & 8.66 & 1.68 & 0.17\\
+5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.86\\
+6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.006\\
+7 & 8.61 & 1.68 & 8.90 & 1.79 & \textless0.001\\
+8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.01\\
 9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
-10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.056\\
+10 & 8.17 & 1.80 & 8.42 & 2.01 & 0.002\\
 \midrule
- & 84.46 & 12.91 & 82.33 & 11.39 & 0.010\\
+ & 84.46 & 12.91 & 82.33 & 11.39 & \textless0.001\\
 \bottomrule
@@
 <<eval=FALSE>>=
+# testItems=function(i, oa, ra){
+#     print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
+# }
 testItems=function(i, oa, ra){
-    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
+    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
 }
@@
 For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
-The short SUS 2.0 consists of 5 items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (leanability, confidence, ease of use, consistency).
+The short SUS 2.0 consists of five items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (confidence, ease of use, consistency, leanability).
 
-[We gotta talk about how this impacts the SUS score calculations, comparison to past scores, etc.]
+The score calculation for the SUS 2.0 is different to that of the original SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale.
+The SUS 2.0 can be scored as a proportion of the maximum score that is achievable.
+First, assign values from 0--4 or 0--6 depending on the number of scale points such that 0 reflects the worse response and 4 or 6 the best.
+Second, sum up the values for all ten responses; the sum should be an integer between 0--48.
+Third, divide by 48 to get the SUS 2.0 score.
+For the short SUS 2.0, follow the same steps except that the sum of response values will lie between 0--24 and you divide by 24, not 48.
 
@@
 
-The two online courses that are compared were offered on two distinct online platforms that shared the same core features but differed considerably in design.
-Based on Nielsen's heuristic evaluation of the two systems, it was determined that one system had more usability problems than the other.
-The following criteria were used [go through these and say which ones were different/better etc.]
+The two systems offered the same basic features, browsing and playing video lectures, but differed considerably in their design.
+We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability.
+While both systems showed generally high usability, one system was deemed superior in these four categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation.
+This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.
 
-Visibility of system status:
-%The system should always keep users informed about what is going on, through appropriate feedback within reasonable time.
-
-Match between system and the real world:
-%The system should speak the user's language, with words, phrases and concepts familiar to the user, rather than system-oriented terms.
-Follow real-world conventions, making information appear in a natural and logical order.
-
-User control and freedom:
-%Users often choose system functions by mistake and will need a clearly marked "emergency exit" to leave the unwanted state without having to go through an extended dialogue. Support undo and redo.
-
-Consistency and standards:
-%Users should not have to wonder whether different words, situations, or actions mean the same thing. Follow platform conventions.
-
-Error prevention:
-%Even better than good error messages is a careful design which prevents a problem from occurring in the first place.
-Either eliminate error-prone conditions or check for them and present users with a confirmation option before they commit to the action.
-
-Recognition rather than recall:
-%Minimize the user's memory load by making objects, actions, and options visible.
-The user should not have to remember information from one part of the dialogue to another.
-Instructions for use of the system should be visible or easily retrievable whenever appropriate.
-
-Flexibility and efficiency of use:
-%Accelerators—unseen by the novice user—may often speed up the interaction for the expert user such that the system can cater to both inexperienced and experienced users.
-Allow users to tailor frequent actions.
-
-Aesthetic and minimalist design:
-%Dialogues should not contain information which is irrelevant or rarely needed.
-Every extra unit of information in a dialogue competes with the relevant units of information and diminishes their relative visibility.
-
-Help users recognize, diagnose, and recover from errors:
-%Error messages should be expressed in plain language (no codes), precisely indicate the problem, and constructively suggest a solution.
-
-Help and documentation:
-%Even though it is better if the system can be used without documentation, it may be necessary to provide help and documentation.
-Any such information should be easy to search, focused on the user's task, list concrete steps to be carried out, and not be too large.
-
-Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for two online course interfaces where one interface has an objectively higher usability level than the other.
+Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for the high-usability and the low-usability system.
 As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.
 While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity.
-A non-parametric Wilcoxon rank sum test for each scale supports this result ($W$=18585 and $p$=0.07 for the original SUS, and $W$=74598 and $p\textless$0.001 for the short SUS 2.0).
+A Mann-Whitney test of the difference between the usability estimates for the two systems for each scale supports this result ($W$=18585, $p$=0.07 for the original SUS; $W$=74598, $p\textless$0.001 for the short SUS 2.0).
 <<eval=FALSE>>=
@@
 
+The paper's unique contribution is to analyze the SUS from a questionnaire design perspective which has not been done before.
+We provide strong evidence that the original SUS induces acquiescence bias and address this and other potential issues by proposing the SUS 2.0: an updated version of the original SUS with statements rephrased as questions with relevant answer scales.
+The SUS 2.0 is less likely to induce response biases, because it conforms to insights from questionnaire design research.
+Moreover, we find that the short SUS 2.0, with only five items, is a more sensitive measure of usability than the original, ten-item SUS.
+
+Future work could investigate the strength of association between the original SUS and the SUS 2.0, which was not possible in these studies as each respondent saw just one usability scale.
+Moreover, our findings are limited in that we could only administer the short SUS 2.0, which constraint us from performing an investigation of the SUS 2.0's factor structure.
+
 \section{Conclusion}
 
+This paper presents compelling evidence that highlights big questionnaire design deficits in the SUS.
+The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a sub-scale of the SUS 2.0 has high reliability, concurrent validity, an single-factor structure, and better sensitivity than the orignal SUS.
+While more work on evaluating the SUS 2.0 is needed, the authors strongly recommend the use of the updated scale for measuring usability.
+
 \section{Acknowledgments}
-We are grateful to the instructors of the two Stanford courses who allowed us conduct survey experiments in their post-course surveys.
+We are grateful to the instructors of the two Stanford courses for allowing us to conduct survey experiments in their post-course surveys.
 
@@
 \centering
-\caption{Items from the proposed SUS 2.0 with answer scales}
+\caption{SUS 2.0 questions with answer scales}
 \label{tab:sus2items}

### 2013-09-14 paper_draft1.Rnw
fixed typos

@@
 \subsection{Heuristic review}
-In this section, we inspect the SUS, one item at a time, for deficiencies in its question design. The item numbers that are refered to correspond to those in Table \ref{tab:items}.
+In this section, we inspect the SUS, one item at a time, for deficiencies in its question design. The item numbers that are referred to correspond to those in Table \ref{tab:items}.
 [Go through each question and describe biases, with references to articles that described them]
@@
 
-As the psychometric propoerties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics is sufficient for this paper.
+As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics is sufficient for this paper.
 Table \ref{tab:dist} provides basic statistics that describe the distribution of the SUS scores.
@@
 Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
-However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant differenece between the original SUS average and reversed SUS reverse-coded average.
+However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.
 
@@
 For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
-The short SUS 2.0 consists of five items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (confidence, ease of use, consistency, leanability).
+The short SUS 2.0 consists of five items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (confidence, ease of use, consistency, learnability).
 
@@
 Third, divide by 48 to get the SUS 2.0 score.
-For the short SUS 2.0, follow the same steps except that the sum of response values will lie between 0--24 and you divide by 24, not 48.
+For the short SUS 2.0, follow the same steps except that the sum of response values will lie between 0--24 and you divide by 24.
 
@@
 We start to establish the SUS 2.0's concurrent validity by investigating its association with related and unrelated constructs.
-We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.22, $t(867)$=7, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.25, $t(867)$=8, $p\textless$0.001); however, the short SUS 2.0 is only marignally associated with the difficulty of the course ($r$=0.08, $t(867)$=2.5, p=0.01).
+We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.22, $t(867)$=7, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.25, $t(867)$=8, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the difficulty of the course ($r$=0.08, $t(867)$=2.5, p=0.01).
 <<eval=FALSE>>=
@@
 Cronbach's $\alpha$ for the short SUS 2.0 is 0.77 based on 869 responses and individual item correlations with the total score vary between 0.47 and 0.86 with 95\% confidence.
-Coefficient alpha is smaller for the short SUS 2.0, but given that coefficient alpha increases with the number of items, the ten-item SUS 2.0 will have higher reliability thant the five-item short SUS 2.0.
+Coefficient alpha is smaller for the short SUS 2.0, but given that coefficient alpha increases with the number of items, the ten-item SUS 2.0 will have higher reliability than the five-item short SUS 2.0.
 <<eval=FALSE>>=
@@
 Future work could investigate the strength of association between the original SUS and the SUS 2.0, which was not possible in these studies as each respondent saw just one usability scale.
-Moreover, our findings are limited in that we could only administer the short SUS 2.0, which constraint us from performing an investigation of the SUS 2.0's factor structure.
+Moreover, our findings are limited in that we could only administer the short SUS 2.0, which constrained us from performing an investigation of the SUS 2.0's factor structure.
 
@@
 This paper presents compelling evidence that highlights big questionnaire design deficits in the SUS.
-The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a sub-scale of the SUS 2.0 has high reliability, concurrent validity, an single-factor structure, and better sensitivity than the orignal SUS.
+The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a subscale of the SUS 2.0 has high reliability, concurrent validity, a single-factor structure, and better sensitivity than the original SUS.
 While more work on evaluating the SUS 2.0 is needed, the authors strongly recommend the use of the updated scale for measuring usability.

### 2013-09-15 paper_draft1.Rnw
edits to evaluation section and scoring

@@
 convSUS = function(x) {
-    return(rowSums((x+1)*2))
+    return(rowSums((x)*2.5))
 }
 convSUS2 = function(x) {
-    return(10*(as.matrix(x+1) %*% (c(5,5,7,7,5)/20))[,1])
-}
-convSUS2s = function(x) {
-    return(10*(as.matrix(x+1) %*% (c(5,7,7,5)/25))[,1])
+    return(rowSums(x)/.18)
 }
@@
 \hypersetup{
-pdftitle={SIGCHI Conference Proceedings Format},
-pdfauthor={LaTeX},
-pdfkeywords={SIGCHI, proceedings, archival format},
+pdftitle={SUS 2.0: Updating the System Usability Scale to Conform with Insights from Recent Questionnaire Design Research},
+pdfauthor={XXXXXXXXX REPLACE WITH AUTHOR NAMES XXXXXXXXXX},
+pdfkeywords={Usability, questionnaire, survey, System Usability Scale, SUS},
 bookmarksnumbered,
@@
 
-\title{SUS 2.0: Updating the System Usability Scale to conform with insights from questionnaire design research}
+\title{SUS 2.0: Updating the System Usability Scale to Conform with Insights from Recent Questionnaire Design Research}
 
@@
 \author{
-  \alignauthor Ren\'{e} F. Kizilcec\\
-    \affaddr{Department of Communication}\\
-    \affaddr{Stanford University}\\
-    \email{kizilcec@stanford.edu}
-  \alignauthor Hendrik Mueller\\
-    \affaddr{UX Research}\\
-    \affaddr{Google}\\
-    \email{hendrikm@google.com}
+    \alignauthor 1st Author Name\\
+    \affaddr{Affiliation}\\
+    \affaddr{Street}\\
+    \affaddr{City}\\
+    \email{E-mail address}
+    \alignauthor 2nd Author Name\\
+    \affaddr{Affiliation}\\
+    \affaddr{Street}\\
+    \affaddr{City}\\
+    \email{E-mail address}
 }
+% \author{
+%   \alignauthor Ren\'{e} F. Kizilcec\\
+%     \affaddr{Department of Communication}\\
+%     \affaddr{Stanford University}\\
+%     \email{kizilcec@stanford.edu}
+%   \alignauthor Hendrik M\"{u}ller\\
+%     \affaddr{UX Research}\\
+%     \affaddr{Google}\\
+%     \email{hendrikm@google.com}
+% }
 
@@
 \begin{abstract}
-The System Usability Scale (SUS) is probably the most widely employed measure of usability.
-Numerous studies have assessed its psychometric properties and used it as a `gold standard' in the development of alternative scales.
-From a questionnaire design perspective, however, the SUS has striking deficiencies which lead to biased measures of usability.
-First, we review the literature on survey biases, inspect each SUS item for issues and show that the SUS is vulnerable to significant acquiescence bias using a survey experiment.
-We then propose the SUS 2.0, an updated version of the SUS, which conforms to insights from questionnaire design research, and present its favorable psychometric properties.
-We present evidence from a second study that suggests that the SUS 2.0 is a more sensitive usability measure than the original SUS.
+The System Usability Scale (SUS), first published in 1996, is probably the most widely employed measure of usability today.
+Numerous studies have assessed its psychometric properties and used it as a “gold standard” in the development of alternative scales.
+Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, have led to the identification of deficiencies for the SUS.
+In this note, we review literature on relevant survey biases, inspect each SUS item for such biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias.
+We then propose the SUS 2.0, an updated version of the SUS, which conforms with recent insights from questionnaire design research, and present its favorable psychometric properties.
+We present evidence from a second study that suggests that the SUS 2.0 is a more sensitive usability measure than the SUS.
 \end{abstract}
 
-\keywords{SUS; questionnaire; surveys}
+\keywords{Usability, questionnaire, survey, System Usability Scale, SUS}
 
@@
 The SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation in the UK.
-It was used as a “quick and dirty” scale to administer after usability studies on electronic office systems, such as the VT100, a text-based terminal system.
+It was used as a ``quick and dirty" scale to administer after usability studies on electronic office systems, such as the VT100, a text-based terminal system.
 SUS measures attitudes and perceptions regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11), yielding a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
@@
 The Likert scale, established by Rensis Likert in 1932, allows questionnaire respondents to specify their level of agreement or disagreement on a symmetric agreement scale for a series of statements \cite{likert1932}.
-SUS’ Likert scale ranges from “Strongly disagree” to “Strongly agree”, with only its endpoints labeled, while additionally all five scale items are numbered from 1 to 5.
-As already noted in Brooke’s initial work \cite{brooke1996}, the phrasing of the statement heavily influences the respondent’s level of agreement or disagreement for it.
+The SUS' Likert scale ranges from ``Strongly disagree" to ``Strongly agree", with only its endpoints labeled, while additionally all five scale items are numbered from 1 to 5.
+As already noted in Brooke’s initial work \cite{brooke1996}, the phrasing of the statement heavily influences the respondent's level of agreement or disagreement for it.
 During the development of the SUS, statements that received the most extreme responses were selected.
@@
 Over the years, SUS has been used across a variety of different systems, including hardware, software, websites, and applications.
-However, since the time of its development in 1986, research regarding questionnaire design has advanced significantly, with several insights that now challenge some of the foundations of the SUS.
-The remainder of this note will explain relevant advances in questionnaire design research, evaluate the original SUS in the context of those, propose an updated version to conform with these insights, and finally offer a comparison between the original and the updated SUS.
+However, since the time of its development in 1986, questionnaire design research has advanced significantly, with several insights that now challenge some of the foundations of the SUS.
+The remainder of this paper will explain relevant advances in questionnaire design research, evaluate the original SUS in the context of those, propose an updated version to conform with these insights, and finally offer a comparison between the original and the updated SUS.
 Note that our intention is not to reduce the number of statements asked about in the SUS, contrary to recent work that attempts to create a shorter usability measure to save time \cite{lewis2013,finstad2010usability}.
-Our goal is to merely update the original questionnaire.
+Our goal is to update the original questionnaire for future use.
 
@@
 
-[Maybe some background on the development and testing of the SUS. Describe the original SUS and a few of its incarnations until today.]
+[NB: Maybe some background on the development and testing of the SUS. Describe the original SUS and a few of its incarnations until today.]
 
@@
 In this section, we inspect the SUS, one item at a time, for deficiencies in its question design. The item numbers that are referred to correspond to those in Table \ref{tab:items}.
-[Go through each question and describe biases, with references to articles that described them]
+
+[NBGo through each question and describe biases, with references to articles that described them]
 
@@
 The survey received 1746 responses.
-At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, their satisfaction with the amount they learnt, and the difficulty of the course.
-Respondents were then randomly assigned to one of three weighted groups: 25\% were presented with the original SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received the SUS 2.0 proposed in this paper (n=869).
+At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, and the difficulty of the course.
+Respondents were then randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received the short SUS 2.0 proposed in this paper (n=869) (see Tables \ref{tab:items} and \ref{tab:sus2items} for scales).
 The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
@@
 
-As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics is sufficient for this paper.
+As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here.
 Table \ref{tab:dist} provides basic statistics that describe the distribution of the SUS scores.
@@
 \centering
-\caption{Statistical information on the SUS distributions (short SUS 2.0 scores scaled to 0-100 for comparison)}
+\caption{Statistical information on the SUS distributions}
 \label{tab:dist}
@@
 \midrule
-Original   & 439 & 38 & 100 & 84.5 & 12.9 & 88 & 16\\
-Reversed   & 438 & 26 & 100 & 82.3 & 11.4 & 84 & 14\\
-Short 2.0  & 869 & 24.5 & 84.5 & 69.7 & 10.4 & 71 & 13\\ 
+Original   & 439 & 22.5 & 100 & 80.6 & 16.1 & 85 & 20\\
+Reversed   & 438 &  7.5 & 100 & 77.9 & 14.2 & 80 & 17.5\\
+Short 2.0  & 869 & 22.2 & 100 & 76.7 & 14.7 & 78 & 22.2\\ 
 \bottomrule
@@
 \end{table}
+<<eval=FALSE>>=
+length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
+length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10]))
+length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5]))
+@
 
 \subsubsection{Concurrent Validity}
-We expect the original SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs.
-We find that the original SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.27, $t(437)$=6, $p\textless$0.001).
-However, there we find no significant association between the SUS and the difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4).
+We expect the  SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs.
+We find that the  SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), and their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001).
+However, there we find no significant association between the SUS and the perceived difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4).
+[NEEDS MORE FRAMING.
+WHY IMPORTANT?]
 <<eval=FALSE>>=
-cor.test(m.orisus$oe,rowSums(m.orisus[,1:10]))
-cor.test(m.orisus$redo,rowSums(m.orisus[,1:10]))
-cor.test(m.orisus$learnsat,rowSums(m.orisus[,1:10]))
-cor.test(m.orisus$difficulty,rowSums(m.orisus[,1:10]))
+cor.test(m.orisus$oe,convSUS2(m.orisus[,1:10]))
+cor.test(m.orisus$redo,convSUS2(m.orisus[,1:10]))
+cor.test(m.orisus$learnsat,convSUS2(m.orisus[,1:10]))
+cor.test(m.orisus$difficulty,convSUS2(m.orisus[,1:10]))
 @
@@
 \subsubsection{Internal consistency}
-For the 439 responses to the original SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence.
-Although the SUS is frequently reported to have a higher coefficient alpha \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an alpha of 0.86 reflects a high degree of interrelatedness.
+For the 439 responses to the SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence.
+Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good degree of interrelatedness.
 <<eval=FALSE>>=
@@
 for(i in 1:10){
-    print(cor.test(rowMeans(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
+    print(cor.test(convSUS2(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
 }
@@
 \subsubsection{Factor analysis}
-A factor analysis of the 439 responses to the original SUS suggests that the scale has a two-factor structure.
+A factor analysis of the 439 responses to the SUS suggests that the scale has a two-factor structure.
 A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one.
@@
 @
-<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for original SUS and SUS 2.0 showing that a single-factor solution is viable, although the original SUS has two factors with eigenvalue greater one'>>=
+<<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for the SUS and short SUS 2.0 showing that the SUS has a two-factor structure, while the short SUS 2.0 has a single-factor structure'>>=
 scree=data.frame(
-    n=c(1:10,1:5), 
-    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,1:5],1)$e.values), 
-    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,5)))
+    n=c(1:10,1:4), 
+    eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,2:5],1)$e.values), 
+    sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,4)))
 ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
@@
 
-\subsection{Acquiescence Bias in SUS}
-A comparison of original SUS scores with scores from the reversed SUS provides strong evidence for acquiescence bias in the SUS.
+\subsection{SUS Induces Acquiescence Bias}
+A comparison of original SUS scores with scores from the reversed SUS provides strong evidence that the SUS induces acquiescence bias.
 Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
@@
 
-Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (this non-parametric alternative of the t-test is used as scores are not normally distributed).
-We find highly significant differences at 99\% significance in all but two items and the overall SUS score.
+Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (a non-parametric alternative of the t-test is used as scores are not normally distributed).
+We find highly significant differences with at least 99\% confidence in all but two items and the overall SUS score.
 This is very strong evidence for the claim that the original SUS induces acquiescence bias.
@@
 \centering
-\caption{Strong evidence for acquiescence bias in the SUS for individual items and the overall score}
+\caption{Means, standard deviations, and p values from Mann-Whitney tests for each item and the overall score of the original and reversed SUS providing strong evidence that the SUS induces acquiescence bias}
 \label{tab:acqui}
@@
 3 & 8.58 & 1.79 & 8.90 & 1.83 & \textless0.001\\
-4 & 8.19 & 2.36 & 8.66 & 1.68 & 0.17\\
-5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.86\\
+4 & 8.19 & 2.36 & 8.66 & 1.68 & 0.173\\
+5 & 8.24 & 1.81 & 8.18 & 1.90 & 0.868\\
 6 & 8.55 & 2.12 & 8.46 & 1.73 & 0.006\\
 7 & 8.61 & 1.68 & 8.90 & 1.79 & \textless0.001\\
-8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.01\\
+8 & 7.99 & 1.85 & 8.19 & 2.02 & 0.013\\
 9 & 8.52 & 2.04 & 7.77 & 2.29 & \textless0.001\\
@@
 \midrule
- & 84.46 & 12.91 & 82.33 & 11.39 & \textless0.001\\
+ & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
 \bottomrule
@@
 
-\section{SUS 2.0 Proposal}
+\section{Proposing the SUS 2.0}
 
-Following the heuristic evaluation of the original SUS items, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias.
+In light of the SUS's deficiencies uncovered in the previous section, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias.
 The first step was to change the questionnaire items from being statements to questions in an effort to reduce acquiescence bias.
-Moreover, statements that were phrased as hypotheticals, such as item 5 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying construct.
+Moreover, statements that were phrased as hypotheticals, such as item 5 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying dimension.
 
-The second step was to change the scale from an agree-disagree Likert scale to scales that reflect the relevant constructs, such as confidence, learnability, or complexity.
-Following question design research, unipolar scales were presented as 5-point scales, while bipolar scales were presented as 7-point scales.
+The second step was to change the scale from an Agree-Disagree Likert scale to scales that reflect the relevant dimension, such as confidence, learnability, or complexity.
+Following recommendations from question design research [CITE], unipolar scales were presented as 5-point scales, while bipolar scales were presented as 7-point scales.
 
-Our goal was to update the SUS items for future use given that the SUS is clearly the most established usability scale.
+Our goal was to update the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale.
 These updated items with corresponding answer scales are presented in Table \ref{tab:sus2items}.
-For our evaluation of the SUS 2.0, we were unable to use the full 10 items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
-The short SUS 2.0 consists of five items marked with asterisks in Table \ref{tab:sus2items} and covers the key facets of the SUS (confidence, ease of use, consistency, learnability).
+For our evaluation of the SUS 2.0, we were unable to use all ten items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
+The short SUS 2.0 consists of four items marked with asterisks in Table \ref{tab:sus2items} and covers the key dimensions of the SUS (confidence, ease of use, consistency, learnability).
 
-The score calculation for the SUS 2.0 is different to that of the original SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale.
-The SUS 2.0 can be scored as a proportion of the maximum score that is achievable.
-First, assign values from 0--4 or 0--6 depending on the number of scale points such that 0 reflects the worse response and 4 or 6 the best.
-Second, sum up the values for all ten responses; the sum should be an integer between 0--48.
-Third, divide by 48 to get the SUS 2.0 score.
-For the short SUS 2.0, follow the same steps except that the sum of response values will lie between 0--24 and you divide by 24.
+The score calculation for the SUS 2.0 is different to that of the SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale.
+To calculate the SUS 2.0 score, first assign values between 0 -- 4 or 0 -- 6 depending on the number of scale points such that 0 reflects the worse usability response and 4 or 6 the best.
+Second, sum up the values for all ten responses to obtain an integer between 0 -- 48.
+Third, divide by 0.48 to obtain the SUS 2.0 score.
+For the short SUS 2.0, follow the same steps except that the sum of response values lies between 0 -- 18 and is divided by 0.18.
+In contrast to the SUS, non-responses are permitted, albeit not encouraged, in the SUS 2.0 and are accounted for by adjusting the denominator accordingly.
 
@@
 
-Table \ref{tab:dist} provides basic statistics on the distribution of short SUS 2.0 scores scaled to the same 0-100 score range as the original SUS to facilitate a comparison.
-Most notably, the SUS 2.0 scores range from about 25 to 85 for a system with good usability, while the original and reversed SUS suffer from ceiling effects which artificially constrain the distribution of scores.
+Table \ref{tab:dist} provides basic statistics on the distribution of short SUS 2.0 scores.
 
 \subsubsection{Concurrent Validity}
-We start to establish the SUS 2.0's concurrent validity by investigating its association with related and unrelated constructs.
-We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001), their likelihood to take another course with the same format ($r$=0.22, $t(867)$=7, $p\textless$0.001), and their satisfaction with how much they learnt ($r$=0.25, $t(867)$=8, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the difficulty of the course ($r$=0.08, $t(867)$=2.5, p=0.01).
+Following the same procedure as for the SUS, we investigate the SUS 2.0's concurrent validity.
+We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001) and their likelihood to take another course with the same format ($r$=0.21, $t(867)$=6, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the perceived difficulty of the course ($r$=0.06, $t(867)$=1.7, p=0.08).
+[FRAMING!]
 <<eval=FALSE>>=
-cor.test(m.sus2$oe,rowSums(m.sus2[,1:5]))
-cor.test(m.sus2$learnsat,rowSums(m.sus2[,1:5]))
-cor.test(m.sus2$redo,rowSums(m.sus2[,1:5]))
-cor.test(m.sus2$difficulty,rowSums(m.sus2[,1:5]))
+cor.test(m.sus2$oe,rowSums(m.sus2[,2:5]))
+cor.test(m.sus2$redo,rowSums(m.sus2[,2:5]))
+cor.test(m.sus2$difficulty,rowSums(m.sus2[,2:5]))
 @
@@
 \subsubsection{Internal consistency}
-Cronbach's $\alpha$ for the short SUS 2.0 is 0.77 based on 869 responses and individual item correlations with the total score vary between 0.47 and 0.86 with 95\% confidence.
-Coefficient alpha is smaller for the short SUS 2.0, but given that coefficient alpha increases with the number of items, the ten-item SUS 2.0 will have higher reliability than the five-item short SUS 2.0.
+Cronbach's $\alpha$ for 869 responses to the short SUS 2.0 is 0.67 and individual item correlations with the total score vary between 0.54 and 0.84 with 95\% confidence.
+Coefficient $\alpha$ is smaller for the short SUS 2.0 than the SUS.
+However, given that coefficient $\alpha$ increases with the number of items, the ten-item SUS 2.0 is expected to have higher reliability than the four-item short SUS 2.0.
 <<eval=FALSE>>=
-alpha(m.sus2[,1:5])$total
+alpha(m.sus2[,2:5])$total
+alpha(a.sus2[,2:10])
+scree(a.sus2[,2:10])
+fa.diagram(fa(a.sus2[,2:10],2))
 omega(m.orisus[,1:10],nfactors=2)
 for(i in 1:5){
-    print(cor.test(rowMeans(m.sus2[,1:5]), m.sus2[,i])$conf[1:2])
+    print(cor.test(convSUS2(m.sus2[,2:5]), m.sus2[,i])$conf[1:2])
 }
@@
 \subsubsection{Factor analysis}
-A factor analysis of 869 responses to the short SUS 2.0 suggests that a single-factor structure is most appropriate for the measure.
-As shown in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point.
+A factor analysis of short SUS 2.0 responses suggests that a single-factor structure is most appropriate for the scale.
+As illustrated in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point.
+A scree test of SUS items that correspond to those in the short SUS 2.0 (items 1, 3, 7, 9) also yields a single-factor solution.
+Hence, the factor structure of the ten-item SUS 2.0 will likely resemble that of the SUS.
 <<eval=FALSE>>=
-fa(m.sus2[,1:5],3)$loadings
-fa(m.sus2[,1:5],2)$loadings
-fa(m.sus2[,1:5],1)$loadings
+scree(m.orisus[,c(1,3,5,7,9)])
+fa(m.sus2[,2:5],3)$loadings
+fa(m.sus2[,2:5],2)$loadings
+fa(m.sus2[,2:5],1)$loadings
 @
@@
 \subsection{Sensitivity of the SUS 2.0}
-A good usability scale should exhibit a high level of sensitivity to reflect subtle differences in usability.
-We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how the SUS 2.0 compares to the original SUS in terms of sensitivity.
+A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability.
+We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how the SUS 2.0 compares to the SUS in terms of sensitivity.
 
-The two systems offered the same basic features, browsing and playing video lectures, but differed considerably in their design.
+The two systems offered the same basic features, i.e.
+browsing and playing video lectures, but differed considerably in their design.
 We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability.
-While both systems showed generally high usability, one system was deemed superior in these four categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation.
+While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation.
 This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.
 
-Figure \ref{fig:sens} illustrates usability ratings using the original SUS and the short SUS 2.0 for the high-usability and the low-usability system.
+Figure \ref{fig:sens} illustrates usability ratings on the SUS and the short SUS 2.0 for the high-usability and the low-usability system.
 As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.
-While the original SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity.
-A Mann-Whitney test of the difference between the usability estimates for the two systems for each scale supports this result ($W$=18585, $p$=0.07 for the original SUS; $W$=74598, $p\textless$0.001 for the short SUS 2.0).
+While the SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the short SUS 2.0 exhibits good sensitivity.
+A Mann-Whitney test of the difference between the usability estimates for the two systems for each scale supports this result ($W$=18585, $p$=0.07 for the SUS; $W$=74598, $p\textless$0.001 for the short SUS 2.0).
 <<eval=FALSE>>=
@@
 
-<<sens, fig.height=3.5, fig.width=4, fig.cap='Evaluation of scale sensitivity for original SUS and SUS 2.0 showing that the SUS 2.0 has high enough sensitivity to distinguish between a high and low usability system'>>=
+<<sens, fig.height=2.5, fig.width=4.5, fig.cap='Evaluation of scale sensitivity for SUS and short SUS 2.0 showing that the latter has high enough sensitivity to distinguish between a high and low usability system, while SUS scores are not significantly different'>>=
 subMean = function(data, indices) {mean(data[indices])}
@@
 a.oriBoot = boot(data=convSUS(a.orisus[,1:10]), statistic=subMean, R=10000)
-m.sus2Boot = boot(data=rowMeans(m.sus2[,2:5]), statistic=subMean, R=10000)
-a.sus2Boot = boot(data=rowMeans(a.sus2[,2:5]), statistic=subMean, R=10000)
+m.sus2Boot = boot(data=rowSums(m.sus2[,2:5])/.18, statistic=subMean, R=10000)
+a.sus2Boot = boot(data=rowSums(a.sus2[,2:5])/.18, statistic=subMean, R=10000)
 
-# boot.ci(m.oriBoot, type="bca") # (83.21, 85.64 )  
-# boot.ci(a.oriBoot, type="bca") # (84.90, 89.29 ) 
-# boot.ci(m.sus2Boot, type="bca") # ( 3.404,  3.494 )
-# boot.ci(a.sus2Boot, type="bca") # ( 3.543,  3.702 )
+# boot.ci(m.oriBoot, type="bca") # (79.03, 82.02 )  
+# boot.ci(a.oriBoot, type="bca") # (81.10, 86.67 ) 
+# boot.ci(m.sus2Boot, type="bca") # (75.67, 77.60 ) 
+# boot.ci(a.sus2Boot, type="bca") # (78.81, 82.33 )
 
 sensDat=data.frame(
-    sus=c("\nOriginal SUS", "\nOriginal SUS", "\nShort SUS 2.0", "\nShort SUS 2.0"),
-    system=c("Low Usability System","High Usability System  ","Low Usability System","High Usability System  "),
+    sus=c("SUS", "SUS", "Short\nSUS 2.0", "Short\nSUS 2.0"),
+    system=c("Low-Usability System","High-Usability System  ","Low-Usability System","High-Usability System  "),
     score=c(m.oriBoot$t0, a.oriBoot$t0, m.sus2Boot$t0, a.sus2Boot$t0),
-    lwr=c(83.21, 84.90, 3.404, 3.543),
-    upr=c(85.64, 89.29, 3.494, 3.702))
+    lwr=c(79.03, 81.10, 75.67, 78.81),
+    upr=c(82.02, 86.67, 77.60, 82.33))
 
@@
     theme_bw() +
-    facet_wrap(~sus, scales="free", nrow=2) +
-    theme(legend.position="bottom", axis.text.y=element_blank(), axis.ticks.y=element_blank(), legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
+    scale_x_continuous(breaks=seq(0,100,2)) +
+    #facet_wrap(~sus, scales="free", nrow=2) +
+    theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
     labs(x="Scores with 95% C.I.", y="", linetype="")
-@
-
-<<eval=FALSE>>=
-length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10]))
-length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10]))
-length(convSUS2(m.sus2[,1:5])); range(convSUS2(m.sus2[,1:5])); mean(convSUS2(m.sus2[,1:5])); sd(convSUS2(m.sus2[,1:5])); median(convSUS2(m.sus2[,1:5])); IQR(convSUS2(m.sus2[,1:5]))
 @
@@
 
-The paper's unique contribution is to analyze the SUS from a questionnaire design perspective which has not been done before.
-We provide strong evidence that the original SUS induces acquiescence bias and address this and other potential issues by proposing the SUS 2.0: an updated version of the original SUS with statements rephrased as questions with relevant answer scales.
+This paper's unique contribution is to analyze the SUS from a questionnaire design perspective which has not been done before.
+We provide strong evidence that the SUS induces acquiescence bias and address this and other potential issues by proposing the SUS 2.0: an updated version of the SUS with statements rephrased as questions with relevant answer scales.
 The SUS 2.0 is less likely to induce response biases, because it conforms to insights from questionnaire design research.
-Moreover, we find that the short SUS 2.0, with only five items, is a more sensitive measure of usability than the original, ten-item SUS.
+Moreover, we find that the short SUS 2.0, with only four items, is a more sensitive measure of usability than the SUS.
+[WRITE ABOUT WHY THIS MIGHT BE]
 
-Future work could investigate the strength of association between the original SUS and the SUS 2.0, which was not possible in these studies as each respondent saw just one usability scale.
-Moreover, our findings are limited in that we could only administer the short SUS 2.0, which constrained us from performing an investigation of the SUS 2.0's factor structure.
+Future work should investigate the strength of association between the SUS and the SUS 2.0, which was not possible in these studies as each respondent saw just one usability scale.
+Moreover, our findings are limited in that we could only administer the short SUS 2.0, which prevented us from performing an investigation of the SUS 2.0's factor structure.
 
@@
 
-This paper presents compelling evidence that highlights big questionnaire design deficits in the SUS.
-The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a subscale of the SUS 2.0 has high reliability, concurrent validity, a single-factor structure, and better sensitivity than the original SUS.
+This paper presents compelling evidence that highlights questionnaire design deficits of the SUS.
+The authors propose a redesign of the scale (the SUS 2.0) and provide evidence that a subscale of the SUS 2.0 has high reliability, concurrent validity, a single-factor structure, and better sensitivity than the SUS.
 While more work on evaluating the SUS 2.0 is needed, the authors strongly recommend the use of the updated scale for measuring usability.
@@
 4  & How cumbersome is it to use the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome \\ \hline
-5$^\star$  & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
+5  & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy \\ \hline
 6  & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex \\ \hline

### 2013-09-15 prep_data.R

@@ -190,27 +190,31 @@ cor(cbind(oe,difficulty,redo,learnsat), use="pair")
 # browser=browser[,1]
 # os=os[,1]
 # res=res[,1]
-
 names(dat)
+
+#How easy or difficult was it to learn how to use the system?
+dat$Q54
+# The scale for this question was displayed incorrectly! Cannot use it!
+
+#How much more is there to learn about the system?
+dat$Q56 #0=nothing at all, 4=a great deal
+q2=4-fac2num(dat$Q56[-1])
+
 #How confident were you using the system?
 dat$Q48 # 0=not at all confident, 4=extremely confident
 q3=fac2num(dat$Q48[-1])
-
-#How consistent is the system?
-dat$Q50 # 0=not at all consistent, 4=extremely consistent
-q5=fac2num(dat$Q50[-1])
 
 #How easy or difficult is it to use the system?
 dat$Q52 #0=extremely diff, 6=extremely easy
 q4=fac2num(dat$Q52[-1])
 
+#How consistent is the system?
+dat$Q50 # 0=not at all consistent, 4=extremely consistent
+q5=fac2num(dat$Q50[-1])
+
 #How likely are you to need support of a technical person to be able to use the system?
 dat$Q60 #0=extremely unlikely, 6=extremely likely
 q6=6-fac2num(dat$Q60[-1])
-
-#How easy or difficult was it to learn how to use the system?
-dat$Q54
-# The scale for this question was displayed incorrectly! Cannot use it!
 
 #How much did you like or dislike the system?
 dat$Q63 #0=extremely dislike, 6=extremely like
@@ -223,10 +227,6 @@ q8=4-fac2num(dat$Q64[-1])
 #How cumbersome is it to use the system?
 dat$Q65 #0=not at all cumbersome, 4=extremely cumbersome
 q9=4-fac2num(dat$Q65[-1])
-
-#How much more is there to learn about the system?
-dat$Q56 #0=nothing at all, 4=a great deal
-q2=4-fac2num(dat$Q56[-1])
 
 #How integrated were the system’s various functions?
 dat$Q61 #0=not at all integrated, 4=extremely integrated

### 2013-09-15 paper_draft1.Rnw
general updates

@@
 %   \alignauthor Hendrik M\"{u}ller\\
-%     \affaddr{UX Research}\\
-%     \affaddr{Google}\\
+%     \affaddr{User Experience Research}\\
+%     \affaddr{Google, Inc.}\\
 %     \email{hendrikm@google.com}
@@
 Numerous studies have assessed its psychometric properties and used it as a “gold standard” in the development of alternative scales.
-Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, have led to the identification of deficiencies for the SUS.
+Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, now challenge some of the foundations of the SUS.
 In this note, we review literature on relevant survey biases, inspect each SUS item for such biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias.
-We then propose the SUS 2.0, an updated version of the SUS, which conforms with recent insights from questionnaire design research, and present its favorable psychometric properties.
+We then propose the SUS 2.0, an updated version of the SUS with neutral questions, which conforms with recent insights from questionnaire design research, and present its favorable psychometric properties.
 We present evidence from a second study that suggests that the SUS 2.0 is a more sensitive usability measure than the SUS.

### 2013-09-15 prep_data.R

@@ -104,16 +104,16 @@ labAD = function(x, rev=0){
 revsus=subset(dat, select=Q75_2:Q75_15)
 revsus[1,]
 rev_sus=data.frame(
-	q1=labAD(revsus[,1]),
-	q2=labAD(revsus[,2]),
-	q3=labAD(revsus[,3],1),
-	q4=labAD(revsus[,4]),
-	q5=labAD(revsus[,5],1),
-	q6=labAD(revsus[,6]),
-	q7=labAD(revsus[,7],1),
-	q8=labAD(revsus[,8],1),
-	q9=labAD(revsus[,9]),
-	q10=labAD(revsus[,10],1))
+	q1=labAD(revsus[,10],1),
+	q2=labAD(revsus[,6]),
+	q3=labAD(revsus[,7],1),
+	q4=labAD(revsus[,2]),
+	q5=labAD(revsus[,8],1),
+	q6=labAD(revsus[,9]),
+	q7=labAD(revsus[,5],1),
+	q8=labAD(revsus[,4]),
+	q9=labAD(revsus[,3],1),
+	q10=labAD(revsus[,1]))
 
 rev_sus_all=data.frame(rev_sus, browser, os, res, oe, difficulty, redo, learnsat)[rowSums(is.na(rev_sus))==0,]
 write.csv(rev_sus_all, "rev_sus_medstat.csv", row.names=F)
@@ -135,16 +135,16 @@ write.csv(rev_sus_all, "rev_sus_medstat.csv", row.names=F)
 orisus=subset(dat, select=Q59_2:Q59_15)
 orisus[1,]
 ori_sus=data.frame(
-	q1=labAD(orisus[,1],1),
-	q2=labAD(orisus[,2],1),
-	q3=labAD(orisus[,3]),
-	q4=labAD(orisus[,4],1),
-	q5=labAD(orisus[,5]),
-	q6=labAD(orisus[,6],1),
-	q7=labAD(orisus[,7]),
-	q8=labAD(orisus[,8]),
-	q9=labAD(orisus[,9],1),
-	q10=labAD(orisus[,10]))
+	q1=labAD(orisus[,10]),
+	q2=labAD(orisus[,6],1),
+	q3=labAD(orisus[,7]),
+	q4=labAD(orisus[,2],1),
+	q5=labAD(orisus[,8]),
+	q6=labAD(orisus[,9],1),
+	q7=labAD(orisus[,5]),
+	q8=labAD(orisus[,4],1),
+	q9=labAD(orisus[,3]),
[…]

### 2013-09-16 paper_draft2.Rnw
added scale image; updated whole draft

\documentclass{sigchi}

\pagenumbering{arabic}

<<setup, include=FALSE, cache=FALSE>>=
opts_chunk$set(fig.path='plots/', fig.align='center', fig.show='asis', echo=FALSE, comment = "", message = FALSE, tidy=FALSE, dev='pdf', out.width="\\linewidth")
options(replace.assign=TRUE)
library(knitr)
library(formatR)
library(ggplot2)
library(grid)
library(boot)
library(lsr)
library(psych)
setwd("~/Dropbox/r-code/SUS2/")
m.sus2 = read.csv("sus2_medstat.csv")
m.orisus = read.csv("ori_sus_medstat.csv")
m.revsus = read.csv("rev_sus_medstat.csv")
a.sus2 = read.csv("sus2_algo.csv")
a.orisus = read.csv("ori_sus_algo.csv")
a.revsus = read.csv("rev_sus_algo.csv")
SE = function(x){
    return(sqrt(var(x)/length(x)))
}
convSUS = function(x) {
    return(rowSums((x)*2.5))
}
convSUS2 = function(x) {
    return(rowSums(x)/.18)
}
@

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
%\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{booktabs}
\usepackage{rotating}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias},
pdfauthor={XXXXXXXXX REPLACE WITH AUTHOR NAMES XXXXXXXXXX},
pdfkeywords={Usability, questionnaire, survey, System Usability Scale, SUS},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{Critiquing the System Usability Scale from a Questionnaire Design Perspective: Beware of Acquiescence Bias}

\numberofauthors{2}
\author{
    \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street}\\
    \affaddr{City}\\
    \email{E-mail address}
    \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Street}\\
    \affaddr{City}\\
    \email{E-mail address}
}
% \author{
%   \alignauthor Ren\'{e} F. Kizilcec\\
%     \affaddr{Department of Communication}\\
%     \affaddr{Stanford University}\\
%     \email{kizilcec@stanford.edu}
%   \alignauthor Hendrik M\"{u}ller\\
%     \affaddr{User Experience Research}\\
%     \affaddr{Google, Inc.}\\
%     \email{hendrikm@google.com}
% }

\maketitle

\begin{abstract}
The System Usability Scale (SUS) is probably the most widely employed measure of usability today. Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scales. Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, now challenge some of the foundations of the SUS. In this note, we review literature on relevant survey biases, inspect each SUS item for such biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias. We then propose a more robust scale, rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
\end{abstract}

\keywords{Usability, questionnaire, survey, System Usability Scale, SUS}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

\section{Introduction}

With the increased focus on developing highly usable products, it has become highly important to actually measure the perceived usability of products or systems. Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement. Commonly used such questionnaires, in order of its first publication, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Software Usability Measurement Inventory (SUMI) \cite{kirakowski1993}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.

Out of all usability measurement questionnaires, the SUS has received the highest level of adoption in both industry and academia, with hundreds of references across publications alone. The SUS is comprised of ten statements measured on a 5-point agreement scale, yielding a single score summarizing the usability assessment of the evaluated system. Over the years, the SUS has been used across a variety of different systems, including hardware, software, websites, applications, and even non-electronic products. However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with several insights that now challenge some of the foundations on which the SUS was developed. Most notably are research insights related to acquiescence bias \cite{smith1967,krosnickPresser2010,sarisKrosnickShaeffer2010}, its relationship to satisficing \cite{krosnick1991,krosnick1999}, and the use of scales with optimal lengths \cite{krosnickFabrigar1997, groves2004}.

The remainder of this note outlines such advances in questionnaire design research relevant to the original SUS, review the SUS in the context of those, propose an alternative version to conform with these insights, and finally compare the original SUS to the proposed alternative. Note that this note's intention is not to reduce the number of statements asked about in the SUS (as attempted by others \cite{finstad2010,lewis2013}), nor the identification and hence elimination of biases that do not relate to acquiescence. Instead, the goal of this note is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.

\section{Related Work}

Even though first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC) in the UK. It was used as a ``quick and dirty" scale to be administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system. The SUS measures attitudes and perceptions regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11). To measure a system's usability on these dimensions, the SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually. The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, which allows questionnaire respondents to specify their level of agreement or disagreement with each of the statements on a symmetric five-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree" (see Figure \ref{fig:originalScale}). Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5. As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statement strongly influences the expressed level of agreement; hence, during the development of the SUS, statements that received the most extreme responses were selected. When analyzing responses to the SUS, the individual responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons. To ensure that this summation is possible, all items of the SUS need to be evaluated by each respondent.

\begin{figure}[!h]
\centering
\includegraphics[width=.9\columnwidth]{originalScale.png}
\caption{5-point agreement scale as used in the SUS}
\label{fig:originalScale}
\end{figure}

One of the questionnaire biases that has been researched thoroughly is that of acquiescence bias, the insight from which are also highly relevant to the SUS. Acquiescence bias is the tendency of a respondent to be more likely to agree with a given statement independent of its substance \cite{smith1967}. Several aspects of questionnaire design contribute to the likelihood for acquiescence bias when responding to a questionnaire. First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than expending cognitive effort to consider reasons for disagreement. This form of shortcutting the question answer process is often referred to as satisficing \cite{krosnick1991}. Second, respondents with lower self-perceived status assume the questionnaire administrator agrees with the posed statement, resulting in deferential agreement bias \cite{sarisKrosnickShaeffer2010}. Similarly, respondents whose personality naturally skews towards agreeableness are more likely to suffer from acquiescence bias \cite{sarisKrosnickShaeffer2010}. Finally, if the respondent's cognitive ability or motivation is lower, acquiescence bias is more likely \cite{krosnickNarayanSmith1996}.

Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}. To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale, instead of using statements with agreement scales \cite{sarisKrosnickShaeffer2010}. Another suggestion to minimize this bias has been to use reverse-keyed constructs, i.e., the same construct is asked positively and negatively in the same survey, the raw scores of which are then combined to correct for acquiescence bias.

Picking the most valid and reliable rating scale for a given question has been another heavily researched topic, especially when moving away from agreement scales. First, research has shown that scale points that are fully labeled as compared to those that just use numbers optimize reliability and minimize bias \cite{groves2004}. Second, the scale length and its items depends on the nature of the construct being measured, i.e., if the construct is unipolar or bipolar in nature. Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, optimizing reliability while minimizing respondent burden \cite{krosnickFabrigar1997}. Bipolar constructs, on the other hand, range from an extreme negative to an extreme positive with a natural midpoint, which are best measured with a 7-point rating scale to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.

\section{Acquiescence in the SUS}

In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias, based on questionnaire design research insights mentioned in the previous section. It also describes an experimental evaluation to identify the effects of acquiescence bias in the SUS. The item numbers of the SUS referred to in the remainder of this note correspond to those in Table \ref{tab:items}.
\subsection{Heuristic review}

Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale. In accordance with questionnaire design research \cite{smith1967}, this particular design encourages the effects of acquiescence bias, i.e., the likelihood for the respondents to simply agree with the statement provided. As an example, let's review item 3 of the SUS: ``I thought the system was easy to use." In this case, respondents are likely to expend unproportional effort on finding reasons that confirm that the system is actually easy to use, instead of thinking about aspects that would contribute to the system being interpreted as difficult to use. As a result, respondents are led towards a more agreeable response in line with ``easy to use" as expressed in the given statement. As another example, let's take a look at item 9 in the original SUS (``I felt very confident using the system"). This statement leads the respondent towards higher confidence, as that's what the statement suggests and since the agreement scale increases the likelihood to agree that the system actually made them feel confident. Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}. As the SUS is often administered at the end of usability studies, hence, not anonymously and with the questionnaire administrator present, this effect may be especially strong. As all other items of the SUS are phrased in the exact same way, this consideration applies equally across all of the SUS.

Notably, some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use"). This suggests that potential acquiescence bias effects may cancel each other out; however, this claim needs further evaluation. Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) continue to suffer from significant acquiescence biases overall. While some of the items (e.g., 4 and 10) are reasonably similar, they are still asking about a slightly different underlying constructs and cannot simply used to cancel out biases.

The SUS' agreement scale is offered as a bipolar scale with 5 items. As agreement/disgreement is a bipolar scale in its nature, the most appropriate scale length should be 7 points, to maximize reliability and data differentiation \cite{krosnickFabrigar1997}. However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths then depend on the nature of the underlying construct being measured, i.e., if the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.

\subsection{Experimental Evaluation of Acquiescence Bias}

\subsubsection{Experiment setup}
Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey. The survey received 1746 responses. 
%At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, and the difficulty of the course. 
Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details). The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos. The rest of the survey was the same for all respondents and contained typical course assessment questions.

\subsubsection{Psychometric Properties of the SUS}

As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here. Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the more robust example scale. Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good level of internal consistency. Moreover, the factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.

\begin{table}[b]
\small
\centering
\begin{tabular}{llll}
\toprule
          & Original & Reversed & Alternative\\
Statistic & SUS      & SUS      & Scale$^\dagger$\\
\midrule
N     & 439 & 438 & 869\\
Range & [23, 100] & [8, 100] & [22, 100]\\
Mean  & 80.6 & 77.9 & 76.7\\
SD    & 16.1 & 14.2 & 14.7\\
Median& 85 & 80 & 78\\
IQR   & 20 & 18 & 22\\
Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
$\left\vert{\lambda}>1\right\vert^{\ast}$ & 2 & 2 & 1\\
\bottomrule
\multicolumn{4}{l}{\scriptsize{$^{\ast}$number of eigenvalues greater than 1 in factor analysis}}\\
\multicolumn{4}{l}{\scriptsize{$^{\dagger}$proposed alternative items 3, 6, 9, 10 from Table \ref{tab:items}}}\\
\end{tabular}
\caption{Statistical and psychometric scale properties.}
\label{tab:dist}
\end{table}
<<eval=FALSE>>=
length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10])); alpha(m.orisus[,1:10])$total; fa(m.orisus[,1:10],1)$e.value
length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10])); alpha(m.revsus[,1:10])$total; fa(m.revsus[,1:10],1)$e.value
length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5])); alpha(m.sus2[,2:5])$total; fa(m.sus2[,2:5],1)$e.value
@

% We expect the  SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs. We find that the  SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), and their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001). However, there we find no significant association between the SUS and the perceived difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4). [NEEDS MORE FRAMING. WHY IMPORTANT?]
% <<eval=FALSE>>=
% cor.test(m.orisus$oe,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$redo,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$learnsat,convSUS2(m.orisus[,1:10]))
% cor.test(m.orisus$difficulty,convSUS2(m.orisus[,1:10]))
% @
% 
% For the 439 responses to the SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence. Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good degree of interrelatedness.
% <<eval=FALSE>>=
% alpha(m.orisus[,1:10])$total
% omega(m.orisus[,1:10], nfactors=2)
% iclust(m.orisus[,1:10],nclusters=2)
% for(i in 1:10){
%     print(cor.test(convSUS2(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
% }
% @
% 
% A factor analysis of the 439 responses to the SUS suggests that the scale has a two-factor structure. A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one. This is consistent with previous work on the SUS factor structure \cite{lewis2009factor}.
% <<eval=FALSE>>=
% fa.parallel(m.orisus[,1:10])
% scree(m.orisus[,1:10])
% fa(m.orisus[,1:10],2)$e.values
% fa.diagram(fa(m.orisus[,1:10],2))
% @
% <<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for the SUS and short SUS 2.0 showing that the SUS has a two-factor structure, while the short SUS 2.0 has a single-factor structure'>>=
% scree=data.frame(
%     n=c(1:10,1:4), 
%     eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,2:5],1)$e.values), 
%     sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,4)))
% ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
% @

\subsubsection{Acquiescence Bias Conclusion}
A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias. Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS. However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.

Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (a non-parametric alternative of the t-test is used as scores are not normally distributed). We find highly significant differences with at least 99\% confidence in all but two items and the overall SUS score. This is very strong evidence for the claim that the original SUS induces acquiescence bias.

\begin{table}[h]
\small
\centering
\begin{tabular}{rccccr}
\toprule
 & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Reversed} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Item \# & M & SD & M & SD & p value \\
\midrule
1 & 3.09 & 0.90 & 3.21 & 1 & 0.002\\
2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
3 & 3.30 & 0.84 & 3.45 & 0.89 & \textless0.001\\
4 & 3.54 & 0.92 & 3.08 & 1.23 & \textless0.001\\
5 & 3.00 & 0.92 & 3.10 & 1.01 & 0.013\\
6 & 3.26 & 1.02 & 2.89 & 1.14 & \textless0.001\\
7 & 3.12 & 0.90 & 3.09 & 0.95 & 0.868\\
8 & 3.10 & 1.18 & 3.33 & 0.84 & 0.173\\
9 & 3.29 & 0.89 & 3.45 & 0.91 & \textless0.001\\
10 & 3.26 & 1.06 & 2.35 & 1.49 & \textless0.001\\
\midrule
overall & 80.58 & 16.14 & 77.92 & 14.23 & \textless0.001\\
\bottomrule
\end{tabular}
\caption{Means, standard deviations, and p values from Mann-Whitney tests for each item and the overall score of the original and reversed SUS providing strong evidence that the SUS induces acquiescence bias.}
\label{tab:acqui}
\end{table}
<<eval=FALSE>>=
# testItems=function(i, oa, ra){
#     print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
# }
testItems=function(i, oa, ra){
    print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(wilcox.test(oa,ra)$p.value,5), sep=" & "))
}
for(i in 1:10){
    testItems(i, m.orisus[,i], m.revsus[,i])
}
testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))

# subMean = function(data, indices) {mean(data[indices])}
# acDat=data.frame(item=rep(1:11, 2), sus=rep(c("Original SUS", "Reversed SUS"),each=11), est=NA, lwr=NA, upr=NA)
# acDat$kind = ifelse(acDat$item==11, "Overall Score", "Individual Item")
# 
# for(i in 1:10){
#     for(s in c("Original SUS", "Reversed SUS")){
#         if(s=="Original SUS"){
#             temp = boot(data=m.orisus[,i], statistic=subMean, R=10000)$t
#         } else {
#             temp = boot(data=m.revsus[,i], statistic=subMean, R=10000)$t
#         }
#      
#         acDat[acDat$item==i & acDat$sus==s, "est"] = mean(temp)
#         acDat[acDat$item==i & acDat$sus==s, "lwr"] = quantile(temp, 0.025)
#         acDat[acDat$item==i & acDat$sus==s, "upr"] = quantile(temp, 0.975)
#     }
# }
# temp = boot(data=convSUS(m.orisus[,1:10]), statistic=subMean, R=10000)
# ci = boot.ci(temp, type="bca")$bca[4:5]
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "est"] = mean(temp$t)
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "lwr"] = ci[1]
# acDat[acDat$item==11 & acDat$sus=="Original SUS", "upr"] = ci[2]
# 
# temp = boot(data=convSUS(m.revsus[,1:10]), statistic=subMean, R=10000)
# ci = boot.ci(temp, type="bca")$bca[4:5]
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "est"] = mean(temp$t)
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "lwr"] = ci[1]
# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "upr"] = ci[2]
# 
# acDat$item = ifelse(acDat$item==11, "all", acDat$item)
#     
# ggplot(acDat, aes(x=est, y=item, xmin=lwr, xmax=upr, linetype=sus)) +
#     geom_point(size=2) +
#     geom_errorbarh(height=.5) +
#     theme_bw() +
#     #scale_x_continuous(breaks=seq(0,100,2)) +
#     facet_wrap(~kind, scales="free", ncol=1) +
#     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
#     labs(x="Scores with 95% C.I.", y="", linetype="")

@

\section{Alternative Scale Proposal}

In light of the SUS's deficiencies uncovered in the previous section, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias, satisficing, and hypothetical projections. The first step was to change the questionnaire items from being statements to questions in an effort to reduce acquiescence bias. Moreover, statements that were phrased as hypotheticals, such as item \#7 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying construct referring back to the system that was just used by the respondent.

The second step was to change the scale from a Likert agreement scale to scales that reflect the relevant construct, such as ease/difficulty (item X), confidence (item X), learnability (item X), or complexity (item X). Following recommendations from question design research \cite{krosnickFabrigar1997}, unipolar constructs were presented as five-point scales, while bipolar constructs were presented as seven-point scales.

Our goal was to update the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale. These updated items with corresponding answer scales are presented in Table \ref{tab:items}. For our evaluation of the SUS 2.0, we were unable to use all ten items and opted for using a reduced number of items which we refer to as the short SUS 2.0. The short SUS 2.0 consists of four items marked with asterisks in Table \ref{tab:items} and covers the key dimensions of the SUS (confidence, ease of use, consistency, learnability).

The score calculation for the SUS 2.0 is different to that of the SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale. To calculate the SUS 2.0 score, first assign values between 0 -- 4 or 0 -- 6 depending on the number of scale points such that 0 reflects the worse usability response and 4 or 6 the best. Second, sum up the values for all ten responses to obtain an integer between 0 -- 48. Third, divide by 0.48 to obtain the SUS 2.0 score. For the short SUS 2.0, follow the same steps except that the sum of response values lies between 0 -- 18 and is divided by 0.18. In contrast to the SUS, non-responses are permitted, albeit not encouraged, in the SUS 2.0 and are accounted for by adjusting the denominator accordingly.

\subsection{Psychometric Properties of Alternative Scale}

This note does not provide a thorough evaluation of the proposed alternative scale. Instead, Table \ref{tab:dist} provides statistical and psychometric information from a subset of items of the alternative scale (items 3, 6, 9, 10) based on 869 responses for comparison with the SUS. Notably, the scales share similar distributional characteristics, but the alternative scale has lower reliability than the SUS and one rather than two-factor structure.

% \subsubsection{Concurrent Validity}
% Following the same procedure as for the SUS, we investigate the SUS 2.0's concurrent validity. We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001) and their likelihood to take another course with the same format ($r$=0.21, $t(867)$=6, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the perceived difficulty of the course ($r$=0.06, $t(867)$=1.7, p=0.08). [FRAMING!]
% <<eval=FALSE>>=
% cor.test(m.sus2$oe,rowSums(m.sus2[,2:5]))
% cor.test(m.sus2$redo,rowSums(m.sus2[,2:5]))
% cor.test(m.sus2$difficulty,rowSums(m.sus2[,2:5]))
% @
% 
% \subsubsection{Internal consistency}
% Cronbach's $\alpha$ for 869 responses to the short SUS 2.0 is 0.67 and individual item correlations with the total score vary between 0.54 and 0.84 with 95\% confidence. Coefficient $\alpha$ is smaller for the short SUS 2.0 than the SUS. However, given that coefficient $\alpha$ increases with the number of items, the ten-item SUS 2.0 is expected to have higher reliability than the four-item short SUS 2.0.
% <<eval=FALSE>>=
% alpha(m.sus2[,2:5])$total
% alpha(a.sus2[,2:10])
% scree(a.sus2[,2:10])
% fa.diagram(fa(a.sus2[,2:10],2))
% omega(m.orisus[,1:10],nfactors=2)
% for(i in 1:5){
%     print(cor.test(convSUS2(m.sus2[,2:5]), m.sus2[,i])$conf[1:2])
% }
% @
% 
% \subsubsection{Factor analysis}
% A factor analysis of short SUS 2.0 responses suggests that a single-factor structure is most appropriate for the scale. As illustrated in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point. A scree test of SUS items that correspond to those in the short SUS 2.0 (items 1, 3, 7, 9) also yields a single-factor solution. Hence, the factor structure of the ten-item SUS 2.0 will likely resemble that of the SUS.
% <<eval=FALSE>>=
% scree(m.orisus[,c(1,3,5,7,9)])
% fa(m.sus2[,2:5],3)$loadings
% fa(m.sus2[,2:5],2)$loadings
% fa(m.sus2[,2:5],1)$loadings
% @

\subsection{Comparing Scale Sensitivity}
A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability. We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how an example of the proposed alternative scale (items 3, 6, 9, 10 from Table \ref{tab:items}) compares to the SUS in terms of sensitivity.

The two systems offered the same basic features, i.e. browsing and playing video lectures, but differed considerably in their design. We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability. While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation. This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.

A Mann-Whitney test of the difference between the usability ratings for the two systems for each scale suggests that one the alternative scale example is more sensitive than the SUS ($W$=18585, $p$=0.069 for the SUS; $W$=17744, $p$=0.014 for the alternative scale). This result is based on 439 usability ratings of the low-usability system and 96 of the high-usability system. Although this comparison uses relatively small sample sizes and only four items from the proposed alternative scale, this finding could point at further potential benefits of using a more robust scale than the SUS.

%Figure \ref{fig:sens} illustrates usability ratings on the SUS and an example of the alternative scale for the high-usability and the low-usability system. As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.  While the SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the alternative scale exhibits good sensitivity. 
<<eval=FALSE>>=
# wilcox.test(convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), conf.int=T)
# wilcox.test(rowSums(m.sus2[,2:5])/.18, rowSums(a.sus2[,2:5])/.18, conf.int=T)

wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]), conf.int=T)
wilcox.test(sample(rowSums(m.sus2[,2:5])/.18,439), sample(rowSums(a.sus2[,2:5])/.18,96), conf.int=T)
@
% 
% <<sens, fig.height=2.5, fig.width=4.5, fig.cap='Evaluation of scale sensitivity for the SUS and alternative scale showing that both are sensitive enough to distinguish between a high and low usability system'>>=
% subMean = function(data, indices) {mean(data[indices])}
% m.oriBoot = boot(data=convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), statistic=subMean, R=10000)
% a.oriBoot = boot(data=convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), statistic=subMean, R=10000)
% m.sus2Boot = boot(data=rowSums(m.sus2[,2:5])/.18, statistic=subMean, R=10000)
% a.sus2Boot = boot(data=rowSums(a.sus2[,2:5])/.18, statistic=subMean, R=10000)
% 
% #  boot.ci(m.oriBoot, type="bca") # (78.21, 80.20 )  
% #  boot.ci(a.oriBoot, type="bca") # (80.96, 85.11 )  
% #  boot.ci(m.sus2Boot, type="bca") # (75.69, 77.63 )  
% #  boot.ci(a.sus2Boot, type="bca") # (78.78, 82.33 )  
% 
% sensDat=data.frame(
%     sus=c("SUS", "SUS", "Alternative\nScale", "Alternative\nScale"),
%     system=c("Low-Usability System","High-Usability System  ","Low-Usability System","High-Usability System  "),
%     score=c(m.oriBoot$t0, a.oriBoot$t0, m.sus2Boot$t0, a.sus2Boot$t0),
%     lwr=c(78.21, 80.96, 75.69, 78.78),
%     upr=c(80.20, 85.11, 77.63, 82.33))
%     
% ggplot(sensDat, aes(x=score, y=sus, xmin=lwr, xmax=upr, linetype=system)) +
%     geom_point(size=2) +
%     geom_errorbarh(height=.5) +
%     theme_bw() +
%     scale_x_continuous(breaks=seq(0,100,2)) +
%     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
%     labs(x="Scores with 95% C.I.", y="", linetype="")
% @

\section{Conclusion}

This note's unique contribution is to critique the SUS from a questionnaire design perspective which has not been done before. We find strong evidence that the SUS induces acquiescence bias and show how a robust alternative scale with statements rephrased as questions and relevant answer scales achieves higher sensitivity in measuring usability than the SUS.

Future work should investigate the strength of association between the SUS and the proposed alternative scale, and establish the latter's reliability, validity, and factor structure. While more work on evaluating the proposed alternative scale is needed, the authors recommend using a more robust scale for measuring usability.

\section{Acknowledgments}
We are grateful to the instructors of the two Stanford courses for allowing us to conduct survey experiments in their post-course surveys.

\begin{sidewaystable}
\small
\centering
\caption{Original and reversed SUS items and updated SUS items with corresponding answer scales}
\label{tab:items}
\begin{tabular}{| r | p{5cm} | p{5cm} | p{3.5cm} | p{6cm} |}
\hline
\textbf{\#}  & \textbf{Original SUS$^\ast$} & \textbf{Reversed SUS$^\ast$} & \textbf{Proposed Alternative} & \textbf{Proposed Answer Scale}\\
\hline
1 & I think that I would like to use this system frequentl & I do not think that I would like to use this system frequently & How much do you like or dislike the system? & \{Extremely, Moderately, Slightly\} dislike, Neither like nor dislike, \{Slightly, Moderately, Extremely\} like \\ \hline
2  & I found the system unnecessarily complex & I found the system appropriately simple & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex\\ \hline
3  & I thought the system was easy to use & I thought the system was hard to use & How easy or difficult is it to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy\\ \hline
4  & I think that I would need the support of a technical person to be able to use this system & I think that I would not need any support of a technical person to be able to use this system  & How likely are you to need the support of a technical person to be able to use the system? & \{Extremely, Very, Somewhat\} unlikely, Neither likely nor unlikely, \{Somewhat, Very, Extremely\} likely\\ \hline
5  & I found the various functions in this system were well integrated & I found the various functions in this system were not well integrated & How integrated are the system's various functions? & \{Not at all, Slightly, Moderately, Very, Extremely\} integrated\\ \hline
6  & I thought there was too much inconsistency in this system & I did not think there was too much inconsistency in this system & How consistent is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} consistent\\ \hline
7  & I would imagine that most people would learn to use this system very quickly  & I would imagine that most people would learn to use this system very slowly & How easy or difficult is it to learn how to use the system? & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy\\ \hline
8  & I found the system very cumbersome to use & I found the system very manageable to use & How cumbersome is it to use the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome\\ \hline
9  & I felt very confident using the system & I did not feel very confident using the system & How confident are you using the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} confident\\ \hline
10  & I needed to learn a lot of things before I could get going with this system   & I needed to learn very few things before I could get going with this system  & How much more is there to learn about the system? & Nothing at all, A little, A moderate amount, A lot, A great deal\\ \hline
\multicolumn{5}{c}{\scriptsize{$^\ast$Items were presented in a matrix with a 5-point Likert scale: Strongly disagree (1), (2), (3), (4), Strongly agree (5)}}
\end{tabular}
\end{sidewaystable}

\bibliographystyle{acm-sigchi}
\bibliography{suslit}
\end{document}

### 2013-09-16 paper_draft3.Rnw
created new shorter draft with all edits from gdoc

--- 2013-09-16 paper_draft2.Rnw
@@
 pdfauthor={XXXXXXXXX REPLACE WITH AUTHOR NAMES XXXXXXXXXX},
-pdfkeywords={Usability, questionnaire, survey, System Usability Scale, SUS},
+pdfkeywords={System Usability Scale, SUS; response biases; usability evaluation; standardized questionnaires},
 bookmarksnumbered,
@@
     \affaddr{Affiliation}\\
-    \affaddr{Street}\\
-    \affaddr{City}\\
+    \affaddr{Street, City}\\
     \email{E-mail address}
@@
 \begin{abstract}
-The System Usability Scale (SUS) is probably the most widely employed measure of usability today.
-Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scales.
-Recent advances in questionnaire design research on satisficing, acquiescence, and other biases, however, now challenge some of the foundations of the SUS.
-In this note, we review literature on relevant survey biases, inspect each SUS item for such biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias.
-We then propose a more robust scale, rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
+The System Usability Scale (SUS) is undoubtedly the most widely employed measure of usability today.
+Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scale proposals.
+Recent advances in questionnaire design research on acquiescence and other biases, however, now challenge some of the foundations of the SUS.
+In this note, we review literature on relevant questionnaire design aspects, inspect the SUS for biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias.
+We then propose an alternative scale, strongly rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
 \end{abstract}
 
-\keywords{Usability, questionnaire, survey, System Usability Scale, SUS}
+\keywords{System Usability Scale, SUS; response biases; usability evaluation; standardized questionnaires}
 
-\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}
+%\terms{Human Factors; Design; Measurement}
+
+\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{User Interfaces–Evaluation/Methodology}
 
@@
 \subsubsection{Experiment setup}
-Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey. The survey received 1746 responses. 
-%At the beginning of the survey respondents were asked to rate their overall experience with the course, their likelihood of taking another course with the same format, and the difficulty of the course.
+Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey.
+The survey received 1746 responses.
 Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details).
@@
 As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here.
-Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the more robust example scale.
-Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good level of internal consistency.
-Moreover, the factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.
+Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the alternative scale proposed below.
+A Cronbach's $\alpha$ of 0.86 reflects a good level of internal consistency, though the SUS is typically reported to have higher reliability \cite{bangor2008,lewis2009factor,borsci2009dimensionality}.
+A quick factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.
 
@@
 Range     & [23, 100] & [8, 100] & [22, 100]\\
-Mean  & 80.6 & 77.9 & 76.7\\
-SD    & 16.1 & 14.2 & 14.7\\
-Median& 85 & 80 & 78\\
-IQR   & 20 & 18 & 22\\
+Mean (SD) & 80.6 (16.1) & 77.9 (14.2) & 76.7 (14.7)\\
+Median (IQR) & 85 (20) & 80 (18) & 78 (22)\\
 Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
@@
 
-% We expect the  SUS to have high concurrent validity in the form of strong associations with related and weak associations with unrelated constructs.
-We find that the  SUS correlates weakly to moderately albeit significantly with the following related constructs: respondents' overall course experience ($r$=0.34, $t(437)$=8, $p\textless$0.001), and their likelihood to take another course with the same format ($r$=0.19, $t(437)$=4, $p\textless$0.001).
-However, there we find no significant association between the SUS and the perceived difficulty of the course ($r$=0.04, $t(437)$=0.9, p=0.4).
-[NEEDS MORE FRAMING.
-WHY IMPORTANT?]
-% <<eval=FALSE>>=
-% cor.test(m.orisus$oe,convSUS2(m.orisus[,1:10]))
-% cor.test(m.orisus$redo,convSUS2(m.orisus[,1:10]))
-% cor.test(m.orisus$learnsat,convSUS2(m.orisus[,1:10]))
-% cor.test(m.orisus$difficulty,convSUS2(m.orisus[,1:10]))
-% @
-% 
-% For the 439 responses to the SUS, Cronbach's $\alpha$ is 0.86 and the correlation of each item with the total score lies between 0.53 and 0.81 with 95\% confidence.
-Although the SUS is frequently reported to have a higher coefficient $\alpha$ \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, an $\alpha$ of 0.86 reflects a good degree of interrelatedness.
-% <<eval=FALSE>>=
-% alpha(m.orisus[,1:10])$total
-% omega(m.orisus[,1:10], nfactors=2)
-% iclust(m.orisus[,1:10],nclusters=2)
-% for(i in 1:10){
-%     print(cor.test(convSUS2(m.orisus[,1:10]), m.orisus[,i])$conf[1:2])
-% }
-% @
-% 
-% A factor analysis of the 439 responses to the SUS suggests that the scale has a two-factor structure.
-A scree plot (Figure \ref{fig:scree}) illustrates that two factors have eigenvalues greater than one.
-This is consistent with previous work on the SUS factor structure \cite{lewis2009factor}.
-% <<eval=FALSE>>=
-% fa.parallel(m.orisus[,1:10])
-% scree(m.orisus[,1:10])
-% fa(m.orisus[,1:10],2)$e.values
-% fa.diagram(fa(m.orisus[,1:10],2))
-% @
-% <<scree, fig.width=4, fig.height=3, fig.cap='Scree plot for the SUS and short SUS 2.0 showing that the SUS has a two-factor structure, while the short SUS 2.0 has a single-factor structure'>>=
-% scree=data.frame(
-%     n=c(1:10,1:4), 
-%     eig=c(fa(m.orisus[,1:10],1)$e.values, fa(m.sus2[,2:5],1)$e.values), 
-%     sus=rep(c("Original SUS", "Short SUS 2.0"),c(10,4)))
-% ggplot(scree, aes(factor(n), eig, group=sus, linetype=sus)) + geom_point() + geom_line()  + geom_hline(yintercept=1) + theme_bw() + labs(x="Factor Number", y="Eigenvalue") + ylim(0,4.5) + theme(legend.title=element_blank(), legend.background=element_blank(), legend.key=element_blank(), legend.key.width=unit(1.5,"cm"), legend.position=c(.65,.75))
-% @
-
-\subsubsection{Acquiescence Bias Conclusion}
+\subsubsection{Acquiescence Bias Result}
 A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias.
@@
 \end{tabular}
-\caption{Means, standard deviations, and p values from Mann-Whitney tests for each item and the overall score of the original and reversed SUS providing strong evidence that the SUS induces acquiescence bias.}
+\caption{Original and reversed SUS scores providing strong evidence for acquiescence bias.}
 \label{tab:acqui}
@@
 <<eval=FALSE>>=
-# testItems=function(i, oa, ra){
-#     print(paste(i, round(mean(oa),2), round(sd(oa),2), round(mean(ra),2), round(sd(ra),2), round(t.test(oa,ra)$p.value,5), sep=" & "))
-# }
 testItems=function(i, oa, ra){
@@
 testItems("all", convSUS(m.orisus[,1:10]), convSUS(m.revsus[,1:10]))
+@
+\section{Alternative Proposal}
 
-# subMean = function(data, indices) {mean(data[indices])}
-# acDat=data.frame(item=rep(1:11, 2), sus=rep(c("Original SUS", "Reversed SUS"),each=11), est=NA, lwr=NA, upr=NA)
-# acDat$kind = ifelse(acDat$item==11, "Overall Score", "Individual Item")
-# 
-# for(i in 1:10){
-#     for(s in c("Original SUS", "Reversed SUS")){
-#         if(s=="Original SUS"){
-#             temp = boot(data=m.orisus[,i], statistic=subMean, R=10000)$t
-#         } else {
-#             temp = boot(data=m.revsus[,i], statistic=subMean, R=10000)$t
-#         }
-#      
-#         acDat[acDat$item==i & acDat$sus==s, "est"] = mean(temp)
-#         acDat[acDat$item==i & acDat$sus==s, "lwr"] = quantile(temp, 0.025)
-#         acDat[acDat$item==i & acDat$sus==s, "upr"] = quantile(temp, 0.975)
-#     }
-# }
-# temp = boot(data=convSUS(m.orisus[,1:10]), statistic=subMean, R=10000)
-# ci = boot.ci(temp, type="bca")$bca[4:5]
-# acDat[acDat$item==11 & acDat$sus=="Original SUS", "est"] = mean(temp$t)
-# acDat[acDat$item==11 & acDat$sus=="Original SUS", "lwr"] = ci[1]
-# acDat[acDat$item==11 & acDat$sus=="Original SUS", "upr"] = ci[2]
-# 
-# temp = boot(data=convSUS(m.revsus[,1:10]), statistic=subMean, R=10000)
-# ci = boot.ci(temp, type="bca")$bca[4:5]
-# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "est"] = mean(temp$t)
-# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "lwr"] = ci[1]
-# acDat[acDat$item==11 & acDat$sus=="Reversed SUS", "upr"] = ci[2]
-# 
-# acDat$item = ifelse(acDat$item==11, "all", acDat$item)
-#     
-# ggplot(acDat, aes(x=est, y=item, xmin=lwr, xmax=upr, linetype=sus)) +
-#     geom_point(size=2) +
-#     geom_errorbarh(height=.5) +
-#     theme_bw() +
-#     #scale_x_continuous(breaks=seq(0,100,2)) +
-#     facet_wrap(~kind, scales="free", ncol=1) +
-#     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
-#     labs(x="Scores with 95% C.I.", y="", linetype="")
+Based on the review and evaluation of the original SUS presented above, this section now discusses a proposal for an alternative scale.
+While reasoning based on recent questionnaire design research is used to change the wording of the different SUS items as well as the response scales, an experiment discussed its quality and sensitivity as compared to SUS.
 
-@
+\subsection{Proposed Wording Changes}
 
-\section{Alternative Scale Proposal}
+To minimize acquiescence bias in the SUS, each of the statements may be transformed into a construct-specific, neutral question with similarly neutral answer options matching the question construct.
+The appropriate scale and its length will then depend on the nature of the construct, i.e., if it is unipolar or bipolar in nature.
+To explain the reasoning that let to the proposed changes, this section now discusses a few exemplary items from the SUS.
+The intention is that the same consideration can then be applied to all other items, as summarized in Table \ref{tab:items}.
 
-In light of the SUS's deficiencies uncovered in the previous section, we propose an updated set of items that, at their core, are equivalent to the SUS, but reduce vulnerability to survey biases, like acquiescence bias, satisficing, and hypothetical projections.
-The first step was to change the questionnaire items from being statements to questions in an effort to reduce acquiescence bias.
-Moreover, statements that were phrased as hypotheticals, such as item \#7 in Table \ref{tab:items}, were rephrased as concrete questions about the underlying construct referring back to the system that was just used by the respondent.
+For example, to minimize acquiescence bias for item 9, we would first need to identify the underlying construct being asked about, which in this case is likely ``confidence".
+The leading statement can then easily be transformed into a construct-specific, neutral question, such as ``How confident were you using the system?".
+To determine the appropriate scale, it is critical to consider the polarity of ``confidence." As confidence naturally starts from a zero point (the absence of confidence), spans to a positive extreme (high confidence), without a natural negative extreme (i.e., several levels of negative confidence), this construct is unipolar in nature.
+As a result, it should then be measured on a 5-point, fully-labeled scale from ``Not at all confident" to ``Extremely confident" (see full scale in Table \ref{tab:items}).
 
-The second step was to change the scale from a Likert agreement scale to scales that reflect the relevant construct, such as ease/difficulty (item X), confidence (item X), learnability (item X), or complexity (item X).
-Following recommendations from question design research \cite{krosnickFabrigar1997}, unipolar constructs were presented as five-point scales, while bipolar constructs were presented as seven-point scales.
+As another example, item 3 refers to the bipolar construct of ease/difficulty.
+Bias may then be minimized by changing its wording to ``How easy or difficult was it to use the system?", giving ``easy" and ``difficult" equal weight so respondents are less led into either direction.
+Due to the bipolarity of this construct, a 7-point, fully labeled scale from ``Extremely difficult" to ``Extremely easy" may be used (see full scale in Table \ref{tab:items}).
+Note, as items 3 and 8 in the original SUS ask about the same underlying construct of ease/difficulty, they would result in the same reworded question, and hence, should only be included once in the full questionnaire.
 
-Our goal was to update the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale.
-These updated items with corresponding answer scales are presented in Table \ref{tab:items}.
-For our evaluation of the SUS 2.0, we were unable to use all ten items and opted for using a reduced number of items which we refer to as the short SUS 2.0.
-The short SUS 2.0 consists of four items marked with asterisks in Table \ref{tab:items} and covers the key dimensions of the SUS (confidence, ease of use, consistency, learnability).
+Our goal was to propose alternative wording for the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale to date.
+Note that the purpose of this note is not an evaluation of the summarization of the questionnaire responses into a single score, hence, it is excluded from this discussion.
 
-The score calculation for the SUS 2.0 is different to that of the SUS, because the SUS 2.0 consists of six items with 5-point unipolar answer scales and four items with 7-point bipolar answer scales, instead of ten items on a 5-point scale.
-To calculate the SUS 2.0 score, first assign values between 0 -- 4 or 0 -- 6 depending on the number of scale points such that 0 reflects the worse usability response and 4 or 6 the best.
-Second, sum up the values for all ten responses to obtain an integer between 0 -- 48.
-Third, divide by 0.48 to obtain the SUS 2.0 score.
-For the short SUS 2.0, follow the same steps except that the sum of response values lies between 0 -- 18 and is divided by 0.18.
-In contrast to the SUS, non-responses are permitted, albeit not encouraged, in the SUS 2.0 and are accounted for by adjusting the denominator accordingly.
+\subsection{Experimental Evaluation}
 
-\subsection{Psychometric Properties of Alternative Scale}
-
-This note does not provide a thorough evaluation of the proposed alternative scale.
-Instead, Table \ref{tab:dist} provides statistical and psychometric information from a subset of items of the alternative scale (items 3, 6, 9, 10) based on 869 responses for comparison with the SUS.
-Notably, the scales share similar distributional characteristics, but the alternative scale has lower reliability than the SUS and one rather than two-factor structure.
-
-% \subsubsection{Concurrent Validity}
-% Following the same procedure as for the SUS, we investigate the SUS 2.0's concurrent validity.
-We find that the short SUS 2.0 correlates moderately and significantly with respondents' overall course experience ($r$=0.31, $t(866)$=10, $p\textless$0.001) and their likelihood to take another course with the same format ($r$=0.21, $t(867)$=6, $p\textless$0.001); however, the short SUS 2.0 is only marginally associated with the perceived difficulty of the course ($r$=0.06, $t(867)$=1.7, p=0.08).
-[FRAMING!]
-% <<eval=FALSE>>=
-% cor.test(m.sus2$oe,rowSums(m.sus2[,2:5]))
-% cor.test(m.sus2$redo,rowSums(m.sus2[,2:5]))
-% cor.test(m.sus2$difficulty,rowSums(m.sus2[,2:5]))
-% @
-% 
-% \subsubsection{Internal consistency}
-% Cronbach's $\alpha$ for 869 responses to the short SUS 2.0 is 0.67 and individual item correlations with the total score vary between 0.54 and 0.84 with 95\% confidence.
-Coefficient $\alpha$ is smaller for the short SUS 2.0 than the SUS.
-However, given that coefficient $\alpha$ increases with the number of items, the ten-item SUS 2.0 is expected to have higher reliability than the four-item short SUS 2.0.
-% <<eval=FALSE>>=
-% alpha(m.sus2[,2:5])$total
-% alpha(a.sus2[,2:10])
-% scree(a.sus2[,2:10])
-% fa.diagram(fa(a.sus2[,2:10],2))
-% omega(m.orisus[,1:10],nfactors=2)
-% for(i in 1:5){
-%     print(cor.test(convSUS2(m.sus2[,2:5]), m.sus2[,i])$conf[1:2])
-% }
-% @
-% 
-% \subsubsection{Factor analysis}
-% A factor analysis of short SUS 2.0 responses suggests that a single-factor structure is most appropriate for the scale.
-As illustrated in the scree plot (Figure \ref{fig:scree}), only one factor has an eigenvalues greater than one and the slope changes considerably at the two factor point.
-A scree test of SUS items that correspond to those in the short SUS 2.0 (items 1, 3, 7, 9) also yields a single-factor solution.
-Hence, the factor structure of the ten-item SUS 2.0 will likely resemble that of the SUS.
-% <<eval=FALSE>>=
-% scree(m.orisus[,c(1,3,5,7,9)])
-% fa(m.sus2[,2:5],3)$loadings
-% fa(m.sus2[,2:5],2)$loadings
-% fa(m.sus2[,2:5],1)$loadings
-% @
-
-\subsection{Comparing Scale Sensitivity}
 A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability.
@@
 While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation.
-This informal usability comparison was the basis for labeling one system as having `high usability' and the other `low usability'.
+This informal usability comparison was the basis for labeling one system as having ``high usability" and the other ``low usability".
+
+\subsubsection{Psychometric Properties of the Alternative Scale}
+
+This note does not provide a thorough evaluation of the proposed alternative scale.
+Instead, Table \ref{tab:dist} provides statistical and psychometric information from a subset of items of the alternative scale (items 3, 6, 9, 10) based on 869 responses for comparison with the SUS.
+Notably, the scales share similar distributional characteristics, but the alternative scale has lower reliability than the SUS and one rather than two-factor structure.
+
+\subsubsection{Scale Sensitivity Result}
 
@@
 
-%Figure \ref{fig:sens} illustrates usability ratings on the SUS and an example of the alternative scale for the high-usability and the low-usability system.
-As the usability scores from both scales were not normally distributed, 95\% confidence intervals were computed from 10,000 bootstrap replicates using the adjusted bootstrap percentile method.
-While the SUS is not sensitive enough to differentiate the usability of the two interfaces with 95\% confidence, the alternative scale exhibits good sensitivity.
 <<eval=FALSE>>=
-# wilcox.test(convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), conf.int=T)
-# wilcox.test(rowSums(m.sus2[,2:5])/.18, rowSums(a.sus2[,2:5])/.18, conf.int=T)
-
 wilcox.test(convSUS(m.orisus[,1:10]), convSUS(a.orisus[,1:10]), conf.int=T)
@@
 @
-% 
-% <<sens, fig.height=2.5, fig.width=4.5, fig.cap='Evaluation of scale sensitivity for the SUS and alternative scale showing that both are sensitive enough to distinguish between a high and low usability system'>>=
-% subMean = function(data, indices) {mean(data[indices])}
-% m.oriBoot = boot(data=convSUS(rbind(m.orisus[,1:10],m.revsus[,1:10])), statistic=subMean, R=10000)
-% a.oriBoot = boot(data=convSUS(rbind(a.orisus[,1:10],a.revsus[,1:10])), statistic=subMean, R=10000)
-% m.sus2Boot = boot(data=rowSums(m.sus2[,2:5])/.18, statistic=subMean, R=10000)
-% a.sus2Boot = boot(data=rowSums(a.sus2[,2:5])/.18, statistic=subMean, R=10000)
-% 
-% #  boot.ci(m.oriBoot, type="bca") # (78.21, 80.20 )  
-% #  boot.ci(a.oriBoot, type="bca") # (80.96, 85.11 )  
-% #  boot.ci(m.sus2Boot, type="bca") # (75.69, 77.63 )  
-% #  boot.ci(a.sus2Boot, type="bca") # (78.78, 82.33 )  
-% 
-% sensDat=data.frame(
-%     sus=c("SUS", "SUS", "Alternative\nScale", "Alternative\nScale"),
-%     system=c("Low-Usability System","High-Usability System  ","Low-Usability System","High-Usability System  "),
-%     score=c(m.oriBoot$t0, a.oriBoot$t0, m.sus2Boot$t0, a.sus2Boot$t0),
-%     lwr=c(78.21, 80.96, 75.69, 78.78),
-%     upr=c(80.20, 85.11, 77.63, 82.33))
-%     
-% ggplot(sensDat, aes(x=score, y=sus, xmin=lwr, xmax=upr, linetype=system)) +
-%     geom_point(size=2) +
-%     geom_errorbarh(height=.5) +
-%     theme_bw() +
-%     scale_x_continuous(breaks=seq(0,100,2)) +
-%     theme(legend.position="bottom", legend.key=element_blank(), strip.background=element_blank(), strip.text=element_text(face="bold"), legend.key.width=unit(1, "cm")) +
-%     labs(x="Scores with 95% C.I.", y="", linetype="")
-% @
 
@@
 This note's unique contribution is to critique the SUS from a questionnaire design perspective which has not been done before.
-We find strong evidence that the SUS induces acquiescence bias and show how a robust alternative scale with statements rephrased as questions and relevant answer scales achieves higher sensitivity in measuring usability than the SUS.
+We find strong evidence that the SUS induces acquiescence bias and show how an alternative scale with statements rephrased as questions and construct-specific, neutral answer scales achieve higher sensitivity in measuring usability than the SUS.
 
-Future work should investigate the strength of association between the SUS and the proposed alternative scale, and establish the latter's reliability, validity, and factor structure.
-While more work on evaluating the proposed alternative scale is needed, the authors recommend using a more robust scale for measuring usability.
+Future work could evaluate the SUS with respect to other biases, such as satisficing and the use of hypotheticals.
+Moreover, a thorough psychometric evaluation of the proposed alternative scale using a variety of systems is needed before we can recommend the adoption of this new scale.
+This will also require addressing the compatibility issue between old and new usability scores for the transition phase.
 
@@
 \centering
-\caption{Original and reversed SUS items and updated SUS items with corresponding answer scales}
+\caption{Original and reversed SUS items and proposed alternative scale with corresponding answer scales.}
 \label{tab:items}
@@
 \hline
-1 & I think that I would like to use this system frequentl & I do not think that I would like to use this system frequently & How much do you like or dislike the system?
+1 & I think that I would like to use this system frequently & I do not think that I would like to use this system frequently & How much do you like or dislike the system?
 & \{Extremely, Moderately, Slightly\} dislike, Neither like nor dislike, \{Slightly, Moderately, Extremely\} like \\ \hline
 2  & I found the system unnecessarily complex & I found the system appropriately simple & How complex is the system? & \{Not at all, Slightly, Moderately, Very, Extremely\} complex\\ \hline
-3  & I thought the system was easy to use & I thought the system was hard to use & How easy or difficult is it to use the system?
+3  & I thought the system was easy to use & I thought the system was hard to use & How easy or difficult was it to use the system?
 & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy\\ \hline
@@
 & \{Not at all, Slightly, Moderately, Very, Extremely\} consistent\\ \hline
-7  & I would imagine that most people would learn to use this system very quickly  & I would imagine that most people would learn to use this system very slowly & How easy or difficult is it to learn how to use the system?
+7  & I would imagine that most people would learn to use this system very quickly  & I would imagine that most people would learn to use this system very slowly & How easy or difficult was it to learn how to use the system?
 & \{Extremely, Moderately, Slightly\} difficult, Neither difficult nor easy, \{Slightly, Moderately, Extremely\} easy\\ \hline
-8  & I found the system very cumbersome to use & I found the system very manageable to use & How cumbersome is it to use the system?
+8  & I found the system very cumbersome to use & I found the system very manageable to use & How cumbersome was it to use the system?
 & \{Not at all, Slightly, Moderately, Very, Extremely\} cumbersome\\ \hline
-9  & I felt very confident using the system & I did not feel very confident using the system & How confident are you using the system?
+9  & I felt very confident using the system & I did not feel very confident using the system & How confident were you using the system?
 & \{Not at all, Slightly, Moderately, Very, Extremely\} confident\\ \hline

### 2013-09-16 paper_draft3.Rnw
phrasing updates from gdoc

@@
 Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
-Commonly used such questionnaires, in order of its first publication, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Software Usability Measurement Inventory (SUMI) \cite{kirakowski1993}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.
+Commonly used such questionnaires, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.
 
@@
 Over the years, the SUS has been used across a variety of different systems, including hardware, software, websites, applications, and even non-electronic products.
-However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with several insights that now challenge some of the foundations on which the SUS was developed.
-Most notably are research insights related to acquiescence bias \cite{smith1967,krosnickPresser2010,sarisKrosnickShaeffer2010}, its relationship to satisficing \cite{krosnick1991,krosnick1999}, and the use of scales with optimal lengths \cite{krosnickFabrigar1997, groves2004}.
+However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with insights that now challenge some of the foundations of the SUS.
+Most notably are insights related to acquiescence bias \cite{smith1967,sarisKrosnickShaeffer2010}, satisficing \cite{krosnick1991,krosnick1999}, and scales lengths \cite{krosnickFabrigar1997, groves2004}.
 
-The remainder of this note outlines such advances in questionnaire design research relevant to the original SUS, review the SUS in the context of those, propose an alternative version to conform with these insights, and finally compare the original SUS to the proposed alternative.
-Note that this note's intention is not to reduce the number of statements asked about in the SUS (as attempted by others \cite{finstad2010,lewis2013}), nor the identification and hence elimination of biases that do not relate to acquiescence.
+The remainder of this note outlines such advances in questionnaire design research relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
+Note that this note's intention is not to reduce the number of statements asked about in the SUS, nor the identification and hence elimination of biases that do not relate to acquiescence.
 Instead, the goal of this note is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.
@@
 
-Even though first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC) in the UK.
-It was used as a ``quick and dirty" scale to be administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system.
-The SUS measures attitudes and perceptions regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11).
-To measure a system's usability on these dimensions, the SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually.
-The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, which allows questionnaire respondents to specify their level of agreement or disagreement with each of the statements on a symmetric five-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree" (see Figure \ref{fig:originalScale}).
-Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5.
-As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statement strongly influences the expressed level of agreement; hence, during the development of the SUS, statements that received the most extreme responses were selected.
-When analyzing responses to the SUS, the individual responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
-To ensure that this summation is possible, all items of the SUS need to be evaluated by each respondent.
+Although first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC).
+It was used as a ``quick and dirty" scale administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system.
+The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11).
+The SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually.
+The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, for respondents to specify their level of agreement or disagreement with each of the statements on a symmetric 5-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree".
+Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5 (see Figure \ref{fig:originalScale}).
+As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statements strongly influences the expressed level of agreement; hence, statements that received the most extreme responses were selected.
+When analyzing responses, the data are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
+To ensure that this summation is possible, all items of the SUS need to be evaluated respondents.
 
@@
 
-One of the questionnaire biases that has been researched thoroughly is that of acquiescence bias, the insight from which are also highly relevant to the SUS.
-Acquiescence bias is the tendency of a respondent to be more likely to agree with a given statement independent of its substance \cite{smith1967}.
+A questionnaire bias that is highly relevant to the SUS and has been researched extensively is acquiescence response bias: the tendency of a respondent to agree with a given statement independent of its substance \cite{smith1967}.
 Several aspects of questionnaire design contribute to the likelihood for acquiescence bias when responding to a questionnaire.
-First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than expending cognitive effort to consider reasons for disagreement.
+First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than affording cognitive effort to consider reasons for disagreement.
 This form of shortcutting the question answer process is often referred to as satisficing \cite{krosnick1991}.
@@
 Similarly, respondents whose personality naturally skews towards agreeableness are more likely to suffer from acquiescence bias \cite{sarisKrosnickShaeffer2010}.
-Finally, if the respondent's cognitive ability or motivation is lower, acquiescence bias is more likely \cite{krosnickNarayanSmith1996}.
+Finally, acquiescence bias is stronger in respondents with lower cognitive ability or motivation \cite{krosnickNarayanSmith1996}.
 
@@
 To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale, instead of using statements with agreement scales \cite{sarisKrosnickShaeffer2010}.
-Another suggestion to minimize this bias has been to use reverse-keyed constructs, i.e., the same construct is asked positively and negatively in the same survey, the raw scores of which are then combined to correct for acquiescence bias.
 
-Picking the most valid and reliable rating scale for a given question has been another heavily researched topic, especially when moving away from agreement scales.
-First, research has shown that scale points that are fully labeled as compared to those that just use numbers optimize reliability and minimize bias \cite{groves2004}.
-Second, the scale length and its items depends on the nature of the construct being measured, i.e., if the construct is unipolar or bipolar in nature.
-Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, optimizing reliability while minimizing respondent burden \cite{krosnickFabrigar1997}.
-Bipolar constructs, on the other hand, range from an extreme negative to an extreme positive with a natural midpoint, which are best measured with a 7-point rating scale to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
+How to pick the most valid and reliable rating scale for a given question has been subject to much research.
+First, research has shown that fully labeled scale points compared to those that just use numbers improve reliability and reduce bias \cite{groves2004}.
+Second, the choice of scale length and labels depends on the nature of the construct being measured, i.e., if it is unipolar or bipolar in nature.
+Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, while bipolar constructs range from an extreme negative to an extreme positive with a natural midpoint and are best measured on a 7-point scale.
+Following these guidelines improves a scale’s reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
 
-\section{Acquiescence in the SUS}
+\section{Acquiescence Bias in the SUS}
 
-In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias, based on questionnaire design research insights mentioned in the previous section.
-It also describes an experimental evaluation to identify the effects of acquiescence bias in the SUS.
-The item numbers of the SUS referred to in the remainder of this note correspond to those in Table \ref{tab:items}.
+In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias. We also describe an experimental evaluation to identify the effects of acquiescence bias in the SUS.
+
 \subsection{Heuristic review}
@@
 Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale.
-In accordance with questionnaire design research \cite{smith1967}, this particular design encourages the effects of acquiescence bias, i.e., the likelihood for the respondents to simply agree with the statement provided.
-As an example, let's review item 3 of the SUS: ``I thought the system was easy to use." In this case, respondents are likely to expend unproportional effort on finding reasons that confirm that the system is actually easy to use, instead of thinking about aspects that would contribute to the system being interpreted as difficult to use.
-As a result, respondents are led towards a more agreeable response in line with ``easy to use" as expressed in the given statement.
-As another example, let's take a look at item 9 in the original SUS (``I felt very confident using the system").
-This statement leads the respondent towards higher confidence, as that's what the statement suggests and since the agreement scale increases the likelihood to agree that the system actually made them feel confident.
+In accordance with questionnaire design research \cite{smith1967}, this particular design induces acquiescence, i.e., an increased likelihood for the respondents to simply agree with the statement provided.
+For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement that the system is easy to use.
+As a result, respondents tend to be more agreeable towards the ``easy to use" description in the statement.
+Consider item 9 from the SUS as another example: ``I felt very confident using the system".
+This loaded phrasing of this statement in combination with the agreement scale induces respondents to report higher levels of confidence than they would otherwise.
+Note also that this item confounds two dimensions by containing ``very", because disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
 Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}.
-As the SUS is often administered at the end of usability studies, hence, not anonymously and with the questionnaire administrator present, this effect may be especially strong.
-As all other items of the SUS are phrased in the exact same way, this consideration applies equally across all of the SUS.
+This effect may be especially strong, as the SUS is frequently administered at the end of usability studies without anonymity and in the presence of the questionnaire administrator.
+These issues apply to all SUS items as they are phrased in the exact same way.
 
 Notably, some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use").
-This suggests that potential acquiescence bias effects may cancel each other out; however, this claim needs further evaluation.
-Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) continue to suffer from significant acquiescence biases overall.
-While some of the items (e.g., 4 and 10) are reasonably similar, they are still asking about a slightly different underlying constructs and cannot simply used to cancel out biases.
+It may be that potential acquiescence bias effects cancel each other out; however, this claim requires empirical evaluation.
+Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) remain vulnerable to significant acquiescence bias.
+While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely mitigating any biases.
 
-The SUS' agreement scale is offered as a bipolar scale with 5 items.
-As agreement/disgreement is a bipolar scale in its nature, the most appropriate scale length should be 7 points, to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
-However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths then depend on the nature of the underlying construct being measured, i.e., if the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.
+The SUS' agreement scale is offered as a bipolar scale with five items.
+As agreement/disagreement is clearly bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
+However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths depend on the nature of the measured underlying construct, i.e.
+whether the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.
 
-\subsection{Experimental Evaluation of Acquiescence Bias}
+\subsection{Experimental Evaluation}
 
-\subsubsection{Experiment setup}
-Participants of a massive open online course offered by Stanford University were asked to complete an optional post-course survey.
-The survey received 1746 responses.
+To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 1746 participants who evaluated an online learning system.
+
+\subsubsection{Experiment Setup}
+
+Participants of a massive open online course were asked to complete an optional post-course survey.
 Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details).
@@
 \midrule
-1 & 3.09 & 0.90 & 3.21 & 1 & 0.002\\
+1 & 3.09 & 0.90 & 3.21 & 1.00 & 0.002\\
 2 & 3.28 & 1.06 & 3.23 & 0.86 & 0.006\\
@@
 \section{Acknowledgments}
-We are grateful to the instructors of the two Stanford courses for allowing us to conduct survey experiments in their post-course surveys.
+We are grateful to the instructors of the two online courses for allowing us to conduct survey experiments in their post-course surveys.
 

### 2013-09-16 paper_draft3.Rnw
shorten sections; merge psychometric properties

@@
 \begin{abstract}
-The System Usability Scale (SUS) is undoubtedly the most widely employed measure of usability today.
-Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scale proposals.
+The System Usability Scale (SUS) is undoubtedly the most widely used measure of usability today.
+Numerous studies have assessed its psychometric properties and used it as a ``gold standard" in the development of alternative scales.
 Recent advances in questionnaire design research on acquiescence and other biases, however, now challenge some of the foundations of the SUS.
-In this note, we review literature on relevant questionnaire design aspects, inspect the SUS for biases, and using a survey experiment, show that the SUS is vulnerable to significant acquiescence bias.
+In this note, we review literature on relevant questionnaire design aspects, inspect the SUS for biases, and using an experiment, show that it is vulnerable to significant acquiescence bias.
 We then propose an alternative scale, strongly rooted in the SUS, which conforms with recent insights from questionnaire design research, and provide an example of how the proposed scale outperforms the SUS in terms of measurement sensitivity.
@@
 \section{Introduction}
+With the increased focus on developing highly usable products, it has become highly important to measure their perceived usability.
+Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
+Commonly used such questionnaires include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, and the System Usability Scale (SUS) \cite{brooke1996}.
 
-With the increased focus on developing highly usable products, it has become highly important to actually measure the perceived usability of products or systems.
-Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
-Commonly used such questionnaires, include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, the Computer System Usability Questionnaires (CSUQ) \cite{lewis1995}, and the System Usability Scale (SUS) \cite{brooke1996}, among several others.
-
-Out of all usability measurement questionnaires, the SUS has received the highest level of adoption in both industry and academia, with hundreds of references across publications alone.
+The SUS has received the highest level of adoption, with hundreds of references across publications alone.
 The SUS is comprised of ten statements measured on a 5-point agreement scale, yielding a single score summarizing the usability assessment of the evaluated system.
-Over the years, the SUS has been used across a variety of different systems, including hardware, software, websites, applications, and even non-electronic products.
-However, since the time of its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with insights that now challenge some of the foundations of the SUS.
+Over the years, the SUS has been used across a variety of systems.
+However, since its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with insights that now challenge some of the foundations of the SUS.
 Most notably are insights related to acquiescence bias \cite{smith1967,sarisKrosnickShaeffer2010}, satisficing \cite{krosnick1991,krosnick1999}, and scales lengths \cite{krosnickFabrigar1997, groves2004}.
 
-The remainder of this note outlines such advances in questionnaire design research relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
+The remainder of this note outlines such advances in questionnaire design relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
 Note that this note's intention is not to reduce the number of statements asked about in the SUS, nor the identification and hence elimination of biases that do not relate to acquiescence.
-Instead, the goal of this note is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.
+Instead, the goal is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.
 
@@
 
-Although first published in 1996, the SUS was developed in 1986 by John Brooke while working at Digital Equipment Corporation (DEC).
+Although first published in 1996, the SUS was developed in 1986 by John Brooke at Digital Equipment Corporation (DEC).
 It was used as a ``quick and dirty" scale administered after usability studies on electronic office systems, such as DEC's VT100, a text-based terminal system.
-The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system (in accordance with the measures of usability defined in ISO 9241-11).
-The SUS is comprised of ten statements (see \ref{tab:items} for their exact wording) which the respondent is asked to rate individually.
-The SUS uses a Likert scale, established by Rensis Likert in 1932 \cite{likert1932}, for respondents to specify their level of agreement or disagreement with each of the statements on a symmetric 5-point agreement scale, ranging from ``Strongly disagree" to ``Strongly agree".
-Note that only its endpoints are labeled, while additionally all five scale items numbered from 1 to 5 (see Figure \ref{fig:originalScale}).
-As already noted in Brooke's initial work \cite{brooke1996}, the phrasing of the statements strongly influences the expressed level of agreement; hence, statements that received the most extreme responses were selected.
-When analyzing responses, the data are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
-To ensure that this summation is possible, all items of the SUS need to be evaluated respondents.
+The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system.
+The SUS is comprised of ten statements (see \ref{tab:items} for their wording) which the respondent is asked to specify their level of agreement or disagreement on a symmetric 5-point Likert scale (as established by Rensis Likert in 1932 \cite{likert1932}).
+Its endpoints are labeled with ``Strongly disagree" and ``Strongly agree", while additionally all five scale items are numbered from 1 to 5 (see Figure \ref{fig:originalScale}).
+All statements of the SUS need to be evaluated by respondents.
+Responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
 
@@
 \centering
-\includegraphics[width=.9\columnwidth]{originalScale.png}
-\caption{5-point agreement scale as used in the SUS}
+\includegraphics[width=.7\columnwidth]{originalScale.png}
+\caption{5-point agreement scale as used in the SUS.}
 \label{fig:originalScale}
@@
 
-A questionnaire bias that is highly relevant to the SUS and has been researched extensively is acquiescence response bias: the tendency of a respondent to agree with a given statement independent of its substance \cite{smith1967}.
-Several aspects of questionnaire design contribute to the likelihood for acquiescence bias when responding to a questionnaire.
-First, when given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than affording cognitive effort to consider reasons for disagreement.
+One of the questionnaire biases researched extensively is that of acquiescence response bias: the tendency of a respondent to agree with a given statement independent of its substance \cite{smith1967}.
+When given a non-neutral statement, respondents are more likely to think of reasons why the statement is true, rather than affording cognitive effort to consider reasons for disagreement.
 This form of shortcutting the question answer process is often referred to as satisficing \cite{krosnick1991}.
-Second, respondents with lower self-perceived status assume the questionnaire administrator agrees with the posed statement, resulting in deferential agreement bias \cite{sarisKrosnickShaeffer2010}.
-Similarly, respondents whose personality naturally skews towards agreeableness are more likely to suffer from acquiescence bias \cite{sarisKrosnickShaeffer2010}.
-Finally, acquiescence bias is stronger in respondents with lower cognitive ability or motivation \cite{krosnickNarayanSmith1996}.
+Furthermore, respondents with lower self-perceived status assume the questionnaire administrator agrees with the posed statement, resulting in deferential agreement bias \cite{sarisKrosnickShaeffer2010}.
+Similarly, respondents whose personality naturally skews towards agreeableness \cite{sarisKrosnickShaeffer2010} and those with lower cognitive ability \cite{krosnickNarayanSmith1996} are more likely to suffer from this bias.
+Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}.
+To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale \cite{sarisKrosnickShaeffer2010}.
 
-Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}.
-To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale, instead of using statements with agreement scales \cite{sarisKrosnickShaeffer2010}.
-
-How to pick the most valid and reliable rating scale for a given question has been subject to much research.
-First, research has shown that fully labeled scale points compared to those that just use numbers improve reliability and reduce bias \cite{groves2004}.
+Determining the current scale and the appropriate length has been another aspect subject to much research.
+First, fully labeled scale points compared to those that just use numbers improve reliability and reduce bias \cite{groves2004}.
 Second, the choice of scale length and labels depends on the nature of the construct being measured, i.e., if it is unipolar or bipolar in nature.
 Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, while bipolar constructs range from an extreme negative to an extreme positive with a natural midpoint and are best measured on a 7-point scale.
-Following these guidelines improves a scale’s reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
+These guidelines improve a scale’s reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
 
@@
 
-In this section, we inspect the original SUS for weaknesses in regards to acquiescence bias. We also describe an experimental evaluation to identify the effects of acquiescence bias in the SUS.
+In this section, we review the SUS in regards to acquiescence bias, first on theoretical grounds and then through empirical evaluation.
 
@@
 Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale.
-In accordance with questionnaire design research \cite{smith1967}, this particular design induces acquiescence, i.e., an increased likelihood for the respondents to simply agree with the statement provided.
-For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement that the system is easy to use.
+In accordance with research \cite{smith1967}, this particular design induces acquiescence.
+For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement.
 As a result, respondents tend to be more agreeable towards the ``easy to use" description in the statement.
 Consider item 9 from the SUS as another example: ``I felt very confident using the system".
-This loaded phrasing of this statement in combination with the agreement scale induces respondents to report higher levels of confidence than they would otherwise.
+The loaded phrasing of this statement in combination with the endpoint-labeled agreement scale induces respondents to report higher levels of confidence than they would otherwise.
 Note also that this item confounds two dimensions by containing ``very", because disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
@@
 Nevertheless, this reverse-keyed approach is used only for a subset of the items in the SUS, hence, all other items (1, 2, 4, 5, 6, 7, 9, 10) remain vulnerable to significant acquiescence bias.
-While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely mitigating any biases.
+While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely to mitigate any biases.
 
 The SUS' agreement scale is offered as a bipolar scale with five items.
-As agreement/disagreement is clearly bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
-However, when transforming the current agreement scale into construct-specific, neutral scales, the different scales and their lengths depend on the nature of the measured underlying construct, i.e.
-whether the construct is unipolar or bipolar in nature \cite{krosnickFabrigar1997}.
+As agreement/disagreement is bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
 
@@
 
-To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 1746 participants who evaluated an online learning system.
-
-\subsubsection{Experiment Setup}
+To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
 
 Participants of a massive open online course were asked to complete an optional post-course survey.
-Respondents were randomly assigned to one of three weighted groups: 25\% were presented with the (original) SUS (n=439), 25\% with the reversed SUS (n=438), and 50\% received an example of a more robust scale proposed in this note (n=869) (see Table \ref{tab:items} for scale details).
-The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
-The rest of the survey was the same for all respondents and contained typical course assessment questions.
+Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (see Table \ref{tab:items} for scale details).
+The system respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
 
-\subsubsection{Psychometric Properties of the SUS}
-
-As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here.
-Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and the alternative scale proposed below.
-A Cronbach's $\alpha$ of 0.86 reflects a good level of internal consistency, though the SUS is typically reported to have higher reliability \cite{bangor2008,lewis2009factor,borsci2009dimensionality}.
-A quick factor analysis yields two eigenvalues greater than one, which is consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.
-
-\begin{table}[b]
-\small
-\centering
-\begin{tabular}{llll}
-\toprule
-          & Original & Reversed & Alternative\\
-Statistic & SUS      & SUS      & Scale$^\dagger$\\
-\midrule
-N         & 439 & 438 & 869\\
-Range     & [23, 100] & [8, 100] & [22, 100]\\
-Mean (SD) & 80.6 (16.1) & 77.9 (14.2) & 76.7 (14.7)\\
-Median (IQR) & 85 (20) & 80 (18) & 78 (22)\\
-Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
-$\left\vert{\lambda}>1\right\vert^{\ast}$ & 2 & 2 & 1\\
-\bottomrule
-\multicolumn{4}{l}{\scriptsize{$^{\ast}$number of eigenvalues greater than 1 in factor analysis}}\\
-\multicolumn{4}{l}{\scriptsize{$^{\dagger}$proposed alternative items 3, 6, 9, 10 from Table \ref{tab:items}}}\\
-\end{tabular}
-\caption{Statistical and psychometric scale properties.}
-\label{tab:dist}
-\end{table}
-<<eval=FALSE>>=
-length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10])); alpha(m.orisus[,1:10])$total; fa(m.orisus[,1:10],1)$e.value
-length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10])); alpha(m.revsus[,1:10])$total; fa(m.revsus[,1:10],1)$e.value
-length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5])); alpha(m.sus2[,2:5])$total; fa(m.sus2[,2:5],1)$e.value
-@
-
-\subsubsection{Acquiescence Bias Result}
 A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias.
@@
 
-\subsection{Experimental Evaluation}
+\subsection{Experimental Evaluation of Scale Sensitivity}
 
 A good usability scale should exhibit a high level of sensitivity to reflect even subtle differences in usability.
-We conducted a second survey study in a post-course survey of an online course that ran on a different system (web interface) to investigate how an example of the proposed alternative scale (items 3, 6, 9, 10 from Table \ref{tab:items}) compares to the SUS in terms of sensitivity.
+We conducted a second survey study in post-course surveys of two online courses that ran on different systems (Web interfaces) to investigate how an example of the proposed alternative scale (items 3, 6, 9, 10 from Table \ref{tab:items}) compares to the SUS in terms of sensitivity.
 
@@
 We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability.
-While both systems showed generally high usability, one system was deemed superior in four evaluation categories: match between system and the real world, consistency and standards, aesthetic and minimalist design, and help and documentation.
+While both systems showed generally high usability, one system was deemed superior in four evaluation categories.
 This informal usability comparison was the basis for labeling one system as having ``high usability" and the other ``low usability".
+We collected 439 alternative scale responses for the low usability system, and 96 for the high usability system.
 
-\subsubsection{Psychometric Properties of the Alternative Scale}
-
-This note does not provide a thorough evaluation of the proposed alternative scale.
-Instead, Table \ref{tab:dist} provides statistical and psychometric information from a subset of items of the alternative scale (items 3, 6, 9, 10) based on 869 responses for comparison with the SUS.
-Notably, the scales share similar distributional characteristics, but the alternative scale has lower reliability than the SUS and one rather than two-factor structure.
-
-\subsubsection{Scale Sensitivity Result}
-
-A Mann-Whitney test of the difference between the usability ratings for the two systems for each scale suggests that one the alternative scale example is more sensitive than the SUS ($W$=18585, $p$=0.069 for the SUS; $W$=17744, $p$=0.014 for the alternative scale).
-This result is based on 439 usability ratings of the low-usability system and 96 of the high-usability system.
-Although this comparison uses relatively small sample sizes and only four items from the proposed alternative scale, this finding could point at further potential benefits of using a more robust scale than the SUS.
-
+A Mann-Whitney test of the difference between the usability ratings for the two systems for each scale suggests that the alternative scale example is more sensitive than the SUS (SUS: $W$=18585, $p$=0.069; alternative scale: $W$=17744, $p$=0.014).
+Although this comparison uses relatively small sample sizes and only four items from the proposed alternative scale, this finding could point at further potential benefits of using a more neutral scale than the SUS.
 <<eval=FALSE>>=
@@
 
+\subsubsection{Comparing Psychometric Properties}
+
+As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here.
+Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and an example (items 3, 6, 9, 10) of the alternative scale proposed above.
+The scales share similar distributional characteristics, but the SUS has higher internal consistency than the alternative scale in terms of Cronbach's $\alpha$.
+However, the SUS is 2.5 times longer than the example of the alternative scale which increases reliability.
+A factor analysis yields two eigenvalues greater than one for the SUS, consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.
+
+\begin{table}[h!]
+\small
+\centering
+\begin{tabular}{llll}
+\toprule
+          & Original & Reversed & Alternative\\
+Statistic & SUS      & SUS      & Scale$^{\text{item \#3,6,9,10}}$\\
+\midrule
+N         & 439 & 438 & 869\\
+Range     & [23, 100] & [8, 100] & [22, 100]\\
+Mean (SD) & 80.6 (16.1) & 77.9 (14.2) & 76.7 (14.7)\\
+Median (IQR) & 85 (20) & 80 (18) & 78 (22)\\
+Cronbach $\alpha$ & 0.86 & 0.73 & 0.67\\
+$\left\vert{\lambda}>1\right\vert^{\ast}$ & 2 & 2 & 1\\
+\bottomrule
+\multicolumn{4}{l}{\scriptsize{$^{\ast}$number of eigenvalues greater than 1 in factor analysis}}\\
+\end{tabular}
+\caption{Statistical and psychometric scale properties.}
+\label{tab:dist}
+\end{table}
+<<eval=FALSE>>=
+length(convSUS(m.orisus[,1:10])); range(convSUS(m.orisus[,1:10])); mean(convSUS(m.orisus[,1:10])); sd(convSUS(m.orisus[,1:10])); median(convSUS(m.orisus[,1:10])); IQR(convSUS(m.orisus[,1:10])); alpha(m.orisus[,1:10])$total; fa(m.orisus[,1:10],1)$e.value
+length(convSUS(m.revsus[,1:10])); range(convSUS(m.revsus[,1:10])); mean(convSUS(m.revsus[,1:10])); sd(convSUS(m.revsus[,1:10])); median(convSUS(m.revsus[,1:10])); IQR(convSUS(m.revsus[,1:10])); alpha(m.revsus[,1:10])$total; fa(m.revsus[,1:10],1)$e.value
+length(convSUS2(m.sus2[,2:5])); range(convSUS2(m.sus2[,2:5])); mean(convSUS2(m.sus2[,2:5])); sd(convSUS2(m.sus2[,2:5])); median(convSUS2(m.sus2[,2:5])); IQR(convSUS2(m.sus2[,2:5])); alpha(m.sus2[,2:5])$total; fa(m.sus2[,2:5],1)$e.value
+@
 \section{Conclusion}

### 2013-09-17 paper_draft3.Rnw
Cliff changes; shortening

@@
 \section{Introduction}
-With the increased focus on developing highly usable products, it has become highly important to measure their perceived usability.
+With the increased focus on developing highly usable products, it has become very important to measure their perceived usability.
 Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
 Commonly used such questionnaires include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, and the System Usability Scale (SUS) \cite{brooke1996}.
-
 The SUS has received the highest level of adoption, with hundreds of references across publications alone.
@@
 However, since its inception in 1986, research regarding the design of valid and reliable questionnaires has advanced significantly, with insights that now challenge some of the foundations of the SUS.
-Most notably are insights related to acquiescence bias \cite{smith1967,sarisKrosnickShaeffer2010}, satisficing \cite{krosnick1991,krosnick1999}, and scales lengths \cite{krosnickFabrigar1997, groves2004}.
+Most relevant are insights related to acquiescence bias \cite{smith1967,sarisKrosnickShaeffer2010}, satisficing \cite{krosnick1991}, and scale lengths \cite{krosnickFabrigar1997, groves2004}.
 
-The remainder of this note outlines such advances in questionnaire design relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
+The remainder of this note outlines advances in questionnaire design relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
 Note that this note's intention is not to reduce the number of statements asked about in the SUS, nor the identification and hence elimination of biases that do not relate to acquiescence.
@@
 The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system.
-The SUS is comprised of ten statements (see \ref{tab:items} for their wording) which the respondent is asked to specify their level of agreement or disagreement on a symmetric 5-point Likert scale (as established by Rensis Likert in 1932 \cite{likert1932}).
+The SUS is comprised of ten statements (see Table \ref{tab:items} for their wording) for which the respondent is asked to specify their level of agreement or disagreement on a symmetric 5-point Likert scale (established by Rensis Likert in 1932).
 Its endpoints are labeled with ``Strongly disagree" and ``Strongly agree", while additionally all five scale items are numbered from 1 to 5 (see Figure \ref{fig:originalScale}).
@@
 Responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=.7\columnwidth]{originalScale.png}
-\caption{5-point agreement scale as used in the SUS.}
-\label{fig:originalScale}
-\end{figure}
 
@@
 Acquiescence bias is the strongest when presented with binary agree/disagree, yes/no, or true/false answer options \cite{smith1967}; however, similar effects have been shown for agreement scales (such as the Likert scale) \cite{sarisKrosnickShaeffer2010}.
-To minimize this bias, research has proposed to refer to the underlying construct in a neutral, non-leading question and offer a neutral scale \cite{sarisKrosnickShaeffer2010}.
+Instead, non-leading questions with neutral scales should be used to measure latent constructs \cite{sarisKrosnickShaeffer2010}.
 
@@
 Unipolar constructs range from zero to an extreme amount and are best measured on a 5-point scale, while bipolar constructs range from an extreme negative to an extreme positive with a natural midpoint and are best measured on a 7-point scale.
-These guidelines improve a scale’s reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
+These guidelines improve a scale's reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.65\columnwidth]{originalScale.png}
+\caption{5-point agreement scale as used in the SUS.}
+\label{fig:originalScale}
+\end{figure}
 
@@
 Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale.
-In accordance with research \cite{smith1967}, this particular design induces acquiescence.
+According to \cite{smith1967}, this particular design induces acquiescence.
 For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement.
@@
 Consider item 9 from the SUS as another example: ``I felt very confident using the system".
-The loaded phrasing of this statement in combination with the endpoint-labeled agreement scale induces respondents to report higher levels of confidence than they would otherwise.
-Note also that this item confounds two dimensions by containing ``very", because disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
+The loaded phrasing of this statement in combination with the endpoint-labeled agreement scale induces respondents to report higher confidence than they would otherwise.
+Note also that this item confounds two dimensions by containing ``very", as disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
 Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}.
-This effect may be especially strong, as the SUS is frequently administered at the end of usability studies without anonymity and in the presence of the questionnaire administrator.
-These issues apply to all SUS items as they are phrased in the exact same way.
+This effect may be especially strong, as the SUS is frequently administered at the end of usability studies without anonymity and in the presence of the researcher.
+These concern applies to all SUS items as they are phrased in the same way.
 
@@
 While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely to mitigate any biases.
-
-The SUS' agreement scale is offered as a bipolar scale with five items.
+The SUS's agreement scale is offered as a bipolar scale with five points.
 As agreement/disagreement is bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
@@
 To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
-
 Participants of a massive open online course were asked to complete an optional post-course survey.
-Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (see Table \ref{tab:items} for scale details).
+Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (Table \ref{tab:items}).
 The system respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
@@
 However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.
-
-Table \ref{tab:acqui} provides means, standard deviations, and p-values from non-parametric Mann-Whitney tests of the hypothesis that there is no location shift (a non-parametric alternative of the t-test is used as scores are not normally distributed).
+Table \ref{tab:acqui} provides means, standard deviations, and p-values from Mann-Whitney tests of the hypothesis that there is no location shift.
 We find highly significant differences with at least 99\% confidence in all but two items and the overall SUS score.
@@
 
-\begin{table}[h]
+\begin{table}
 \small
@@
 
-Based on the review and evaluation of the original SUS presented above, this section now discusses a proposal for an alternative scale.
-While reasoning based on recent questionnaire design research is used to change the wording of the different SUS items as well as the response scales, an experiment discussed its quality and sensitivity as compared to SUS.
+In this section we propose and discuss an alternative usability scale.
+Recent questionnaire design insights are used to change the wording of the different SUS items and the response scales.
+An experiment illustrates its quality and sensitivity as compared to SUS.
 
@@
 
-To minimize acquiescence bias in the SUS, each of the statements may be transformed into a construct-specific, neutral question with similarly neutral answer options matching the question construct.
+To minimize acquiescence bias, each of the SUS statements may be transformed into a construct-specific, neutral question with similarly neutral answer options matching the question construct.
 The appropriate scale and its length will then depend on the nature of the construct, i.e., if it is unipolar or bipolar in nature.
-To explain the reasoning that let to the proposed changes, this section now discusses a few exemplary items from the SUS.
-The intention is that the same consideration can then be applied to all other items, as summarized in Table \ref{tab:items}.
+To explain the reasoning that led to the proposed changes, let us consider a few exemplary items from the SUS; however, with the intention that the same can be applied to all other items, as summarized in Table \ref{tab:items}.
 
-For example, to minimize acquiescence bias for item 9, we would first need to identify the underlying construct being asked about, which in this case is likely ``confidence".
-The leading statement can then easily be transformed into a construct-specific, neutral question, such as ``How confident were you using the system?".
-To determine the appropriate scale, it is critical to consider the polarity of ``confidence." As confidence naturally starts from a zero point (the absence of confidence), spans to a positive extreme (high confidence), without a natural negative extreme (i.e., several levels of negative confidence), this construct is unipolar in nature.
-As a result, it should then be measured on a 5-point, fully-labeled scale from ``Not at all confident" to ``Extremely confident" (see full scale in Table \ref{tab:items}).
+For example, for item 9, we would first need to identify the underlying construct being asked about, which in this case is likely ``confidence".
+The leading statement can then easily be transformed into a construct-specific, neutral question: ``How confident were you using the system?".
+To determine the appropriate scale, it is critical to consider the polarity of ``confidence." As confidence naturally starts from a zero point (the absence of confidence), spans to a positive extreme (high confidence), without a natural negative extreme (levels of negative confidence), this construct is unipolar in nature.
+As a result, it should be measured on a 5-point, fully-labeled scale from ``Not at all confident" to ``Extremely confident" (see full scale in Table \ref{tab:items}).
 
-As another example, item 3 refers to the bipolar construct of ease/difficulty.
+As another example, item 3 refers to the bipolar construct of ``ease/difficulty".
 Bias may then be minimized by changing its wording to ``How easy or difficult was it to use the system?", giving ``easy" and ``difficult" equal weight so respondents are less led into either direction.
 Due to the bipolarity of this construct, a 7-point, fully labeled scale from ``Extremely difficult" to ``Extremely easy" may be used (see full scale in Table \ref{tab:items}).
-Note, as items 3 and 8 in the original SUS ask about the same underlying construct of ease/difficulty, they would result in the same reworded question, and hence, should only be included once in the full questionnaire.
-
-Our goal was to propose alternative wording for the SUS items instead of creating a new measure, given that the SUS is probably the most established usability scale to date.
-Note that the purpose of this note is not an evaluation of the summarization of the questionnaire responses into a single score, hence, it is excluded from this discussion.
+Note, as items 3 and 8 in the original SUS ask about the same construct, only one of these identical reworded questions would be included in the questionnaire.
+Note that the purpose of this note is not to evaluate the summarization of the questionnaire responses into a single score, hence, it is excluded from this discussion.
 
@@
 
-The two systems offered the same basic features, i.e.
-browsing and playing video lectures, but differed considerably in their design.
+The two systems offered similar capabilities, i.e., browsing and playing video lectures, but differed considerably in their design.
 We employed Molich and Nielsen's heuristic evaluation criteria \cite{molich1990improving} to informally establish which system has better usability.
@@
 
-A Mann-Whitney test of the difference between the usability ratings for the two systems for each scale suggests that the alternative scale example is more sensitive than the SUS (SUS: $W$=18585, $p$=0.069; alternative scale: $W$=17744, $p$=0.014).
-Although this comparison uses relatively small sample sizes and only four items from the proposed alternative scale, this finding could point at further potential benefits of using a more neutral scale than the SUS.
+A Mann-Whitney test of the difference between the usability ratings for the two systems for each scale suggests that the alternative scale is more sensitive than the SUS ($W$=18585, $p$=0.069 for the SUS; $W$=17744, $p$=0.014 for the alternative scale).
+The alternative scale scores are significantly different for the two systems, while the SUS scores are not.
+Although this comparison uses relatively small sample sizes and only four items from the proposed alternative scale, this finding points at further potential benefits of using a more neutral scale than the SUS.
 <<eval=FALSE>>=
@@
 \subsubsection{Comparing Psychometric Properties}
-
-As the psychometric properties of the SUS have been studied extensively \cite{bangor2008,lewis2009factor,borsci2009dimensionality}, a brief evaluation of key statistics should be sufficient here.
+%As the psychometric properties of the SUS have been studied extensively \cite{lewis2009factor}, a brief evaluation of key statistics should be sufficient here.
 Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and an example (items 3, 6, 9, 10) of the alternative scale proposed above.
@@
 We find strong evidence that the SUS induces acquiescence bias and show how an alternative scale with statements rephrased as questions and construct-specific, neutral answer scales achieve higher sensitivity in measuring usability than the SUS.
-
 Future work could evaluate the SUS with respect to other biases, such as satisficing and the use of hypotheticals.
@@
 This will also require addressing the compatibility issue between old and new usability scores for the transition phase.
-
-\section{Acknowledgments}
-We are grateful to the instructors of the two online courses for allowing us to conduct survey experiments in their post-course surveys.
 

### 2013-09-17 paper_draft3.Rnw
updated references and shortened psychometric prop

@@
 
-\subsection{Experimental Evaluation}
-
-To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
-Participants of a massive open online course were asked to complete an optional post-course survey.
-Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (Table \ref{tab:items}).
-The system respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
-
-A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias.
-Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
-However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.
-Table \ref{tab:acqui} provides means, standard deviations, and p-values from Mann-Whitney tests of the hypothesis that there is no location shift.
-We find highly significant differences with at least 99\% confidence in all but two items and the overall SUS score.
-This is very strong evidence for the claim that the original SUS induces acquiescence bias.
-
 \begin{table}
@@
 \end{table}
+
+\subsection{Experimental Evaluation}
+
+To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
+Participants of a massive open online course were asked to complete an optional post-course survey.
+Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (Table \ref{tab:items}).
+The system respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
+
+A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias.
+Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
+However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.
+Table \ref{tab:acqui} provides means, standard deviations, and p-values from Mann-Whitney tests of the hypothesis that there is no location shift.
+We find highly significant differences with at least 99\% confidence in all but two items and the overall SUS score.
+This is very strong evidence for the claim that the original SUS induces acquiescence bias.
+
+\section{Alternative Proposal}
+
+In this section we propose and discuss an alternative usability scale.
+Recent questionnaire design insights are used to change the wording of the different SUS items and the response scales.
+An experiment illustrates its quality and sensitivity as compared to SUS.
+
 <<eval=FALSE>>=
@@
 @
-\section{Alternative Proposal}
-
-In this section we propose and discuss an alternative usability scale.
-Recent questionnaire design insights are used to change the wording of the different SUS items and the response scales.
-An experiment illustrates its quality and sensitivity as compared to SUS.
 
@@
 %As the psychometric properties of the SUS have been studied extensively \cite{lewis2009factor}, a brief evaluation of key statistics should be sufficient here.
-Table \ref{tab:dist} provides an overview of statistics and psychometric properties of the original and reversed SUS, and an example (items 3, 6, 9, 10) of the alternative scale proposed above.
-The scales share similar distributional characteristics, but the SUS has higher internal consistency than the alternative scale in terms of Cronbach's $\alpha$.
-However, the SUS is 2.5 times longer than the example of the alternative scale which increases reliability.
-A factor analysis yields two eigenvalues greater than one for the SUS, consistent with previous work on the SUS's factor structure \cite{lewis2009factor}.
+Table \ref{tab:dist} provides statistics and psychometric properties of the original and reversed SUS, and an example (items 3, 6, 9, 10) of the alternative scale.
+They share similar distributional characteristics, but the SUS has higher reliability than the alternative scale in terms of Cronbach's $\alpha$.
+A factor analysis yields two eigenvalues greater than one for the SUS, consistent with previous work on its factor structure \cite{lewis2009factor}.
 

### 2013-09-17 paper_draft3.Rnw
shortened intro

@@
 With the increased focus on developing highly usable products, it has become very important to measure their perceived usability.
-Especially during usability studies, large or small, standardized questionnaires are being used widely for such measurement.
-Commonly used such questionnaires include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, and the System Usability Scale (SUS) \cite{brooke1996}.
+Standardized questionnaires are widely used for such measurement in small and large usability studies.
+Commonly used questionnaires include the Computer User Satisfaction Inventory (CUSI) \cite{kirakowski1987}, the Questionnaire for User Interface Satisfaction (QUIS) \cite{chin1988}, the After Scenario Questionnaire (ASQ) \cite{lewis1991}, and the System Usability Scale (SUS) \cite{brooke1996}.
 The SUS has received the highest level of adoption, with hundreds of references across publications alone.

### 2013-09-18 paper_draft3.Rnw
made final wording changes and moved fig/tab to better place

@@
 The remainder of this note outlines advances in questionnaire design relevant to the SUS, reviews the original SUS in the context of those, proposes an alternative version to conform with these insights, and finally compares the SUS to the proposed alternative.
-Note that this note's intention is not to reduce the number of statements asked about in the SUS, nor the identification and hence elimination of biases that do not relate to acquiescence.
+Note that the authors' intention is not to reduce the number of items in the SUS, nor the identification and subsequent elimination of biases unrelated to acquiescence.
 Instead, the goal is to critique the wording of the statements and response options of the SUS on theoretical grounds with empirical evidence and propose an alternative scale.
@@
 The SUS measures attitudes regarding the effectiveness, efficiency, and satisfaction with a system.
-The SUS is comprised of ten statements (see Table \ref{tab:items} for their wording) for which the respondent is asked to specify their level of agreement or disagreement on a symmetric 5-point Likert scale (established by Rensis Likert in 1932).
+The SUS is comprised of ten statements (see Table \ref{tab:items}) for which the respondent is asked to specify their level of agreement or disagreement on a symmetric 5-point Likert scale (established by Rensis Likert in 1932).
 Its endpoints are labeled with ``Strongly disagree" and ``Strongly agree", while additionally all five scale items are numbered from 1 to 5 (see Figure \ref{fig:originalScale}).
@@
 Responses are consolidated into a single score to represent the global usability assessment for that system and to enable cross-system comparisons.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.65\columnwidth]{originalScale.png}
+\caption{5-point agreement scale as used in the SUS.}
+\label{fig:originalScale}
+\end{figure}
 
@@
 
-Determining the current scale and the appropriate length has been another aspect subject to much research.
+The design of answer scales with optimal labeling and length has been subject to much research.
 First, fully labeled scale points compared to those that just use numbers improve reliability and reduce bias \cite{groves2004}.
@@
 These guidelines improve a scale's reliability and maximizes data differentiation while minimizing the cognitive burden on respondents \cite{krosnickFabrigar1997}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=.65\columnwidth]{originalScale.png}
-\caption{5-point agreement scale as used in the SUS.}
-\label{fig:originalScale}
-\end{figure}
 
@@
 
-Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale.
-According to \cite{smith1967}, this particular design induces acquiescence.
-For instance, consider item 3 of the SUS (see Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement.
+Each of the ten items of the SUS is constructed of a non-neutral statement and an agreement scale, which is a design that has been shown to induce acquiescence \cite{smith1967}.
+For instance, consider item 3 of the SUS (Table \ref{tab:items}): ``I thought the system was easy to use." Respondents are likely to spend relatively more effort on finding reasons that confirm rather than contradict the statement.
 As a result, respondents tend to be more agreeable towards the ``easy to use" description in the statement.
-Consider item 9 from the SUS as another example: ``I felt very confident using the system".
+As another example, consider item 9 from the SUS: ``I felt very confident using the system".
 The loaded phrasing of this statement in combination with the endpoint-labeled agreement scale induces respondents to report higher confidence than they would otherwise.
-Note also that this item confounds two dimensions by containing ``very", as disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
+Note also that this item confounds two dimensions by containing ``very" and ``confident", as disagreement with feeling very confident is not equivalent to feeling unconfident, but could theoretically indicate extreme confidence.
 Additionally, the nature of the agreement scale leads respondents further towards agreement as disagreeing with anyone requires courage and cognitive effort \cite{sarisKrosnickShaeffer2010}.
 This effect may be especially strong, as the SUS is frequently administered at the end of usability studies without anonymity and in the presence of the researcher.
-These concern applies to all SUS items as they are phrased in the same way.
+These concerns apply to all SUS items as they are phrased in the same way.
 
-Notably, some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use").
+Some of the items in the SUS have been reverse-keyed to ask about the same construct from opposite directions, as for example in items 3 (``easy to use") and 8 (``cumbersome to use").
 It may be that potential acquiescence bias effects cancel each other out; however, this claim requires empirical evaluation.
@@
 While some items (e.g., 4 and 10) are reasonably similar, they effectively tap into different underlying constructs and are thus unlikely to mitigate any biases.
-The SUS's agreement scale is offered as a bipolar scale with five points.
+Lastly, the SUS's agreement scale is offered as a bipolar scale with five points.
 As agreement/disagreement is bipolar, the appropriate scale length would be seven points to maximize reliability and data differentiation \cite{krosnickFabrigar1997}.
 
-\begin{table}
+\subsection{Experimental Evaluation}
+
+To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
+Participants of a massive open online course were asked to complete an optional post-course survey.
+Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (Table \ref{tab:items}).
+The system that respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
+
+\begin{table}[h!]
 \small
@@
 \cmidrule(r){2-3} \cmidrule(r){4-5}
-Item \# & M & SD & M & SD & p value \\
+Item \# & M & SD & M & SD & p-value \\
 \midrule
@@
 
-\subsection{Experimental Evaluation}
-
-To provide empirical support for the principal claim that the SUS is vulnerable to acquiescence bias in particular, we ran an online survey experiment with 877 participants who evaluated an online learning system.
-Participants of a massive open online course were asked to complete an optional post-course survey.
-Respondents were randomly assigned to one of two groups: one received the original SUS (n=439) and the other the reversed SUS (n=438) (Table \ref{tab:items}).
-The system respondents were asked to evaluate comprised of the course sites for browsing and watching lecture videos.
-
 A comparison between scores from the original and reversed SUS provides strong evidence that the SUS induces acquiescence bias.
 Without acquiescence bias, the average for each item on the original SUS would not be significantly different from the reverse-coded average for each item on the reversed SUS.
-However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in a significant difference between the original SUS average and reversed SUS reverse-coded average.
+However, if acquiescence bias exists, respondents would tend to agree with statements independent of the statement's tone, which would be reflected in significant differences between the original SUS scores and reversed SUS reverse-coded scores.
 Table \ref{tab:acqui} provides means, standard deviations, and p-values from Mann-Whitney tests of the hypothesis that there is no location shift.
@@
 The appropriate scale and its length will then depend on the nature of the construct, i.e., if it is unipolar or bipolar in nature.
-To explain the reasoning that led to the proposed changes, let us consider a few exemplary items from the SUS; however, with the intention that the same can be applied to all other items, as summarized in Table \ref{tab:items}.
+To explain the reasoning that led to the proposed changes, we consider a few exemplary items from the SUS with the intention that the same logic can be applied to the remaining items, as summarized in Table \ref{tab:items}.
 
@@
 Note, as items 3 and 8 in the original SUS ask about the same construct, only one of these identical reworded questions would be included in the questionnaire.
-Note that the purpose of this note is not to evaluate the summarization of the questionnaire responses into a single score, hence, it is excluded from this discussion.
+As the purpose of this note is not to evaluate the summarization of the questionnaire responses into a single score, it is excluded from this discussion.
 
@@
 This informal usability comparison was the basis for labeling one system as having ``high usability" and the other ``low usability".
-We collected 439 alternative scale responses for the low usability system, and 96 for the high usability system.
+We collected 439 SUS and alternative scale responses for the low usability system, and 96 for the high usability system.
 
@@
 They share similar distributional characteristics, but the SUS has higher reliability than the alternative scale in terms of Cronbach's $\alpha$.
+The SUS is also 2.5 times longer than the alternative scale example which increases reliability.
 A factor analysis yields two eigenvalues greater than one for the SUS, consistent with previous work on its factor structure \cite{lewis2009factor}.
@@
 \centering
-\caption{Original and reversed SUS items and proposed alternative scale with corresponding answer scales.}
+\caption{Original and reversed SUS items (see Figure \ref{fig:originalScale} for answer scale) and proposed alternative with corresponding answer scales.}
 \label{tab:items}
@@
 & Nothing at all, A little, A moderate amount, A lot, A great deal\\ \hline
-\multicolumn{5}{c}{\scriptsize{$^\ast$Items were presented in a matrix with a 5-point Likert scale: Strongly disagree (1), (2), (3), (4), Strongly agree (5)}}
 \end{tabular}
